% =============================================================================
% Severity-Aware Evaluation of Large Language Models:
% An Empirical Study with Actuarial Frequency-Severity Models
%
% Companion empirical paper to NeurIPS 2026 position paper:
%   "Machine Learning Evaluation Should Adopt Actuarial Frequency-Severity Model"
% =============================================================================

\documentclass[11pt]{article}

% ACL / EMNLP style
\usepackage[review,hyperref]{acl}
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{inconsolata}
\usepackage{csquotes}

% Packages
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{enumitem}
\setlist{nosep}
\setlist[itemize]{leftmargin=*,label={}}
\usepackage{url}
\usepackage{colortbl}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{pgfplots}
\pgfplotsset{compat=1.17}
% Consistent severity palette across all figures
% Colorblind-safe palette (Wong 2011, "Color blindness", Nature Methods 8:441).
% Avoids the red/green pair entirely. sevNeg = blue (good), sevCrit =
% vermillion (bad), sevMaj = orange, sevMin = sky blue. accentBlue is
% reserved for figures that show a single quantitative variable.
\definecolor{sevNeg}{HTML}{0072B2}
\definecolor{sevMin}{HTML}{56B4E9}
\definecolor{sevMaj}{HTML}{E69F00}
\definecolor{sevCrit}{HTML}{D55E00}
\definecolor{accentBlue}{HTML}{0072B2}

% Capitalise \autoref output to match ACL style.
\renewcommand{\sectionautorefname}{Section}
\renewcommand{\subsectionautorefname}{Section}
\renewcommand{\subsubsectionautorefname}{Section}
\renewcommand{\figureautorefname}{Figure}
\renewcommand{\tableautorefname}{Table}

% QFrCoRE-style prompt template (Beauchemin et al. 2025) -- two stacked
% rounded TikZ nodes using the exact colours from the reference paper:
% tgb1 = rgb(0.2666,0.466,0.666) for system, tgb4 = rgb(0.8,0.733,0.2666)
% for user.
\definecolor{tgb1}{rgb}{0.2666, 0.466, 0.666}
\definecolor{tgb4}{rgb}{0.8, 0.733, 0.2666}
\newcommand{\promptbox}[2]{%
  \par\smallskip\noindent
  \begin{tikzpicture}[every node/.style={transform shape}]
    \node[rectangle, rounded corners, draw=tgb1, fill=tgb1!80,
          text width=0.92\linewidth, align=left, inner sep=1ex]
        (sysbox) {%
          $\ll$system$\gg$ \ttfamily #1};
    \node[rectangle, rounded corners, draw=tgb4, fill=tgb4,
          below=0.1cm of sysbox,
          text width=0.92\linewidth, align=left, inner sep=1ex]
        (userbox) {%
          $\ll$user$\gg$ \ttfamily #2};
  \end{tikzpicture}\par\smallskip
}

% Custom commands
\newcommand{\ES}{\mathbb{E}[S]}
\newcommand{\VaR}{\mathrm{VaR}}
\newcommand{\TVaR}{\mathrm{TVaR}}
\newcommand{\seveval}{\texttt{severity-eval}}
\newcommand{\bpi}{\boldsymbol{\pi}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}

\title{Severity-Aware Evaluation of Large Language Models:\\
An Empirical Study with Actuarial Frequency-Severity Models}

\author{
  David Beauchemin\textsuperscript{1} \quad
  Richard Khoury\textsuperscript{1} \quad
  Vincent Goulet\textsuperscript{2} \\
  \textsuperscript{1}Department of Computer Science and Software Engineering, Universit\'{e} Laval \\
  \textsuperscript{2}\'{E}cole d'actuariat, Universit\'{e} Laval \\
  \texttt{\{david.beauchemin,richard.khoury\}@ulaval.ca} \\
  \texttt{vincent.goulet@act.ulaval.ca}
}

\begin{document}
\maketitle

% ============================================================================
% ABSTRACT
% ============================================================================

\begin{abstract}
Accuracy treats every LLM error as equally costly. The actuarial
frequency-severity framework of \citet{beauchemin2026position}
addresses this in theory but was demonstrated on a single
$82$-query insurance corpus. We provide the first multi-domain
empirical evaluation: $12$ open-weight LLMs ($8$\,B--$120$\,B
parameters) on $11$ benchmarks across finance, medicine, law,
and insurance, with severity labels derived from each corpus's
structured fields. At full corpus, accuracy and $\widehat{\ES}$
rankings are positively correlated in every domain (Kendall
$\tau$ from $0.58$ in finance to $0.97$ in legal); the
small-sample divergence reproduces in finance and insurance at
$n{=}1000$ per cell but not at full scale. An MQM-weighted
baseline tracks our ranking with $\tau$ from $0.82$ to $1.00$,
so the framework's value lies less in the ordering than in the
interpretable dollar magnitudes: $74$ accuracy-vs-cost
inversions (median $\$1.9$\,M, max $\$28.7$\,M per $10{,}000$
outputs), a variance decomposition whose severity share moves
from $0.05$ (medical) to $0.81$ (finance) with the cross-model
$p$/$\mu_X$ coefficient-of-variation ratio, a routing tradeoff
that cuts $\widehat{\ES}$ by $85$\,\% in finance at $18$\,\%
deferral and by $92$--$98$\,\% in the other three domains at
$41$--$53$\,\% deferral, and per-domain Lundberg reserves from
$\$0.8$\,M to $\$6.2$\,M per $10{,}000$ outputs per period. We
release \seveval{}, a HuggingFace-style Python library, with
annotations and calibration evidence.
\end{abstract}

% ============================================================================
% 1. INTRODUCTION
% ============================================================================

\section{Introduction}
\label{sec:introduction}

Accuracy treats every error as equally costly: a misclassified
spam email and a misclassified cancer diagnosis count the same.
As LLMs enter high-stakes deployments, that assumption breaks.
On JudgeBERT (French legal simplification), DeepSeek-R1-70B
edges Llama-3.3-70B by $+0.7$ accuracy points yet costs
$\$28.7$\,M more per $10{,}000$ outputs at calibrated severity
weights; the leaderboard rewards the more expensive model. A
revenue answer of \$$11.6$\,B instead of \$$14.2$\,B shifts a
trading decision, and a missed sepsis cue delays intervention.
Per-error costs span orders of magnitude, yet accuracy assigns
them the same penalty.

\citet{beauchemin2026position} proposed importing actuarial
frequency-severity models from insurance mathematics. The
framework splits operational risk into an error rate $p$ and a
severity distribution $\bpi$ over cost levels, yielding a
compound loss $S = \sum_{i=1}^{N} X_i$ whose expected value, tail
quantiles, and required reserves are computable in closed form
or by Monte Carlo. That contribution was conceptual: feasibility
on a single $82$-query insurance corpus, no validation across
domains, models, or full-scale corpora; the dramatic
accuracy-vs-cost ranking divergence reported there had unknown
stability at scale. We provide that validation: $12$ open-weight LLMs
($8$\,B--$120$\,B parameters, dense, MoE, and reasoning-tuned)
on $11$ benchmarks across finance, medicine, law, and insurance,
$\sim$$440$\,K predictions over $132$ (model, dataset) cells
each evaluated on the \emph{full} test set. The aim is not to
crown a best model; it is to characterise where the framework
adds signal beyond accuracy, and where it does not.

Our three contributions:

\begin{enumerate}[leftmargin=*]
  \item \textbf{Rankings agree at scale; the framework's value
    is in magnitudes.} At full corpus, per-domain Kendall $\tau$
    between accuracy and $\ES$ rankings is $0.58$--$0.97$ (all
    $p < 0.01$); the dramatic small-sample divergence reproduces
    at $n{=}1000$ per cell in finance and insurance but not at
    full scale, and never in medical or legal. An MQM-weighted
    baseline tracks our ranking with per-domain $\tau$ from
    $0.82$ to $1.00$. The framework's contribution is therefore
    not the ordering but the dollar magnitudes: $74$ inversions
    ($\$0.32$\,B mis-attributed) with median gap $\$1.9$\,M and
    max $\$28.7$\,M, plus tail measures and reserves that no
    weighted-mean baseline can produce.
  \item \textbf{The mechanism behind \emph{when} severity
    matters.} Inter-model variance in $\ES$ splits into a
    severity share that ranges from $5\%$ (medical) to $81\%$
    (finance); the split is driven by which input has the larger
    cross-model coefficient of variation ($p$-CV vs.\
    $\mu_X$-CV). This makes a domain a priori classifiable as
    severity- or frequency-dominated, telling a procurement
    committee whether $\bpi$ needs inspection.
  \item \textbf{A routing tradeoff with two operating regimes
    and per-domain reserves.} Modelled as excess-of-loss
    reinsurance, HITL routing cuts $\ES$ by $85\%$ at $18\%$
    deferral in finance (cheap routing) and by $92$--$98\%$ at
    $41$--$53\%$ deferral in the other three domains (deep
    routing). Tail premium $\TVaR_{0.95}/\ES$ ranges from
    $1.06$ in legal to $1.25$ in insurance; Lundberg reserves
    ($\psi^\ast{=}0.01$, $\theta{=}0.2$) range from $\$0.8$\,M
    (finance) to $\$6.2$\,M (medical) per $10{,}000$ outputs
    per period.
\end{enumerate}

\noindent We release \seveval{}, a HuggingFace-style Python
library (\autoref{sec:app-library}), the rule-based severity
tables (\autoref{sec:severity-annotation}), and the cost
calibrations (\autoref{sec:app-calibration}).

% ============================================================================
% 2. RELATED WORK
% ============================================================================

\section{Related Work}
\label{sec:related-work}

\paragraph{Evaluation metrics and cost-sensitive learning.}
Accuracy ignores error heterogeneity. NLP metrics that partly
address this (BLEU \citep{papineni2002bleu}, ROUGE
\citep{lin2004rouge}, BERTScore \citep{zhang2020bertscore},
BLEURT \citep{sellam2020bleurt}) measure linguistic quality, not
operational risk; MQM \citep{lommel2014mqm} weights minor/major/
critical errors at $1/5/25$, the closest NLP precedent. Cost-sensitive
learning \citep{elkan2001foundations,drummond2006cost,ling2008cost}
optimises models under a known cost matrix at training time but
is restricted to classification with fixed classes. None of these
maps errors to dollar-denominated organisational losses.

\paragraph{Risk-aware AI evaluation and actuarial loss models.}
Regulatory frameworks (EU AI Act \citep{euaiact2024}, NIST AI RMF
\citep{nist2023rmf}, ISO/IEC~42001 \citep{iso42001}) prescribe
risk tiering without quantitative loss measures. Risk taxonomies
\citep{weidinger2022taxonomy,bergman2024stela,kirk2024personalising}
and ontologies \citep{golpayegani2022airo} formalise harms
qualitatively. The Basel~II/III operational-risk Loss
Distribution Approach \citep{basel2006} and its cyber-risk
extensions \citep{malavasi2022cyber,eling2016cyber} use the same
frequency-severity decomposition we adopt; recent Federal Reserve
work \citep{bosfed2025ai} applies it to AI-related operational
losses. \citet{beauchemin2026position} bridge actuarial methods
and AI evaluation; our paper is the empirical companion.

\paragraph{HITL deferral and large-scale LLM benchmarks.}
Routing uncertain predictions to a human has been studied under
\emph{learning to defer}
\citep{madras2018predict,mozannar2020consistent}, \emph{selective
prediction} \citep{elyaniv2010selective,geifman2017selective}, and
post-hoc deferral estimators
\citep{narasimhan2022posthoc,okati2021differentiable}; we recast
it as excess-of-loss reinsurance. Large-scale LLM evaluation
suites (HELM \citep{liang2023helm}, MMLU
\citep{hendrycks2021mmlu}, Open LLM Leaderboard
\citep{beeching2023openllm}, Chatbot Arena
\citep{chiang2024chatbot}) and domain-specific benchmarks
\citep{islam2023financebench,chen2021finqa,jin2021medqa,kazi2024medcalc,hendrycks2021cuad,koreeda2021contractnli}
all report accuracy-class metrics. Our work augments these
benchmarks with a per-instance severity label and the actuarial
risk measures that follow.

% ============================================================================
% 3. FRAMEWORK
% ============================================================================

\section{Actuarial Framework}
\label{sec:framework}

We recapitulate the framework of \citet{beauchemin2026position};
formal derivations are in the original paper.

\subsection{Frequency-Severity Decomposition}

We follow the standard collective-risk model of actuarial science
\citep{klugman2012loss}: an AI system produces $n$ outputs per
evaluation period, each correct (probability $1-p$) or erroneous
(probability $p$). Each error falls into one of $K$ severity
categories with costs $c_1 < \cdots < c_K$ and probabilities
$\bpi = (\pi_1, \ldots, \pi_K)$:
\begin{align}
  N &\sim \mathrm{Bin}(n, p), \label{eq:frequency} \\
  X_i &\in \{c_1, \ldots, c_K\}, \quad \mathbb{P}(X_i = c_k) = \pi_k. \label{eq:severity}
\end{align}

\subsection{Compound Loss}

The aggregate loss $S = \sum_{i=1}^{N} X_i$ has closed-form moments:
\begin{align}
  \ES &= n p \, \mu_X, &
  \mathrm{Var}(S) &= n p \, \sigma_X^2 + n p (1\!-\!p) \, \mu_X^2,
  \label{eq:expected_loss}
\end{align}
where $\mu_X = \sum_k c_k \pi_k$ and
$\sigma_X^2 = \sum_k c_k^2 \pi_k - \mu_X^2$. The expected loss
$\ES$ captures both error frequency ($p$) and economic impact
($\mu_X$): two systems with identical accuracy can differ arbitrarily
in $\ES$ if $\bpi_A \neq \bpi_B$.

\paragraph{What the cost model quantifies.}
The severity tier of an instance reflects the cost of
\emph{erring on that question} given its structured field (clause
type, deal-point category, drug interaction type, etc.), not the
cost of the specific wrong output a model produces. For
extractive and classification tasks (CUAD, MAUD, ContractNLI,
HEAD-QA, MedMCQA, DDI, PrivacyQA, JudgeBERT) the two coincide:
missing an indemnification clause has the same cost regardless
of what the model outputs instead. For the three generative
finance benchmarks (FinanceBench, FinQA, TAT-QA) the
question-level cost is a conservative upper bound: a $\$11.6$\,B
revenue answer instead of $\$14.2$\,B is operationally severe,
while $\$14.0$\,B is wrong but operationally benign. Per-output
severity scoring is left to future work (see Limitations).

\subsection{Risk Measures}

We compute the distribution of $S$ via Monte Carlo simulation
($10^5$ replications) and extract two tail risk measures: the
\textbf{Value-at-Risk}
$\VaR_\alpha(S) = \inf\{s : \mathbb{P}(S \leq s) \geq \alpha\}$,
i.e.\ the loss exceeded with probability $1 - \alpha$; and the
\textbf{Tail Value-at-Risk}
$\TVaR_\alpha(S) = \mathbb{E}[S \mid S > \VaR_\alpha(S)]$,
the expected loss in the worst $(1 - \alpha)$ scenarios.

\subsection{Ruin Theory and Reserves}

For multi-period deployment, the surplus process
$U_t = u + nrt - \sum_{s=1}^{t} S_s$ tracks the remaining risk
budget, where $u$ is the initial reserve and $r$ is the net revenue
per output. The ruin probability
$\psi(u) = \mathbb{P}(\exists\, t \geq 1 : U_t < 0)$ is bounded
by Lundberg's inequality $\psi(u) \leq e^{-Ru}$, where $R$ is the
adjustment coefficient obtained by solving
$\lambda(M_X(R) - 1) = nrR$ with $\lambda = np$
\citep{dickson2016insurance}. This yields a minimum reserve
$u^* = -\ln(\psi^*) / R$ for a target ruin tolerance $\psi^*$.

\subsection{Human-in-the-Loop Routing}

Following the excess-of-loss reinsurance analogy
\citep{beauchemin2026position}, outputs with estimated severity above
a retention threshold $d$ are routed to human review at cost $h$.
The routing ratio $\rho = \sum_{k : c_k > d} \pi_k$ gives the
fraction redirected, yielding total expected cost:
\begin{equation}
  \mathbb{E}[C] = n p (1 - \rho) \, \mu_X^{\mathrm{ret}} + n p \rho \, h,
  \label{eq:routing}
\end{equation}
where $\mu_X^{\mathrm{ret}}$ is the mean severity of retained errors.

% ============================================================================
% 4. EXPERIMENTAL SETUP
% ============================================================================

\section{Experimental Setup}
\label{sec:experiments}

\subsection{Datasets}
\label{sec:datasets}

We use $11$ benchmarks across four high-stakes domains
(\autoref{tab:datasets}). The selection criterion is that
severity must be derivable from a \emph{structured} field of the
source dataset, or from a defensible matrix over structured
answer/program/metric axes (finance benchmarks only). Three
datasets initially considered (MedQA, MedCalc-Bench, and the RAG
Insurance corpus) were dropped after an inter-annotator audit
could not defend their keyword-based labels at the
critical-vs-major boundary (\autoref{sec:limitations}). Per-domain
severity mapping rules and tier-by-tier worked examples are in
\autoref{sec:app-severity-by-dataset}.

\begin{table*}[t]
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}llrlll@{}}
\toprule
Domain & Dataset & Inst.\ & Task & Severity source & Reference \\
\midrule
\multirow{3}{*}{Finance}
  & FinanceBench  & 150        & Extr.\ QA          & Answer $\times$ metric matrix       & \citet{islam2023financebench} \\
  & FinQA         & 1{,}147    & Num.\ reasoning    & Answer $\times$ metric matrix       & \citet{chen2021finqa} \\
  & TAT-QA        & 1{,}663    & Table+text QA      & Answer $\times$ metric matrix       & \citet{zhu2021tatqa} \\
\midrule
\multirow{3}{*}{Medical}
  & HEAD-QA       & 12{,}751   & MCQ                & Academic category                   & \citet{vilares2019headqa} \\
  & MedMCQA       & 182{,}822  & MCQ                & Specialty (\texttt{subject\_name})  & \citet{pal2022medmcqa} \\
  & DDI~2013      & 34{,}449   & Relation extr.   & Relation type (effect/advice)       & \citet{herrerozazo2013ddi} \\
\midrule
\multirow{3}{*}{Legal}
  & CUAD          & 4{,}182    & Clause extr.\      & Clause type ($41$)                  & \citet{hendrycks2021cuad} \\
  & MAUD          & 39{,}231   & Merger MCQ         & Deal-point category                 & \citet{wang2023maud} \\
  & ContractNLI   & 9{,}788    & NLI                & Fixed hypothesis                    & \citet{koreeda2021contractnli} \\
\midrule
\multirow{2}{*}{Insurance}
  & JudgeBERT     & 1{,}483    & Simplification     & Clause category $\times$ score      & \citet{beauchemin2025thesis} \\
  & PrivacyQA     & 157{,}420  & Relevance     & OPP-115 category mapping            & \citet{ravichander2019privacyqa} \\
\bottomrule
\end{tabular}
\caption{Datasets evaluated. \enquote{Inst.} denotes the number of
test-set instances.}
\label{tab:datasets}
\end{table*}

% ----------------------------------------------------------------------------
\subsection{Severity Annotation}
\label{sec:severity-annotation}

Each domain uses a four-level severity taxonomy ($K = 4$) with cost
levels calibrated against empirical evidence. \autoref{tab:costs}
summarizes the cost vectors.

\begin{table}[t]
\centering
\small
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrr}
\toprule
Domain & $c_1$ (negl.) & $c_2$ (minor) & $c_3$ (major) & $c_4$ (crit.) \\
\midrule
Finance   & \$100  & \$1k   & \$10k  & \$100k \\
Medical   & \$500  & \$5k   & \$50k  & \$500k \\
Legal     & \$200  & \$2k   & \$20k  & \$200k \\
Insurance & \$100  & \$2k   & \$10k  & \$250k \\
\bottomrule
\end{tabular}}
\caption{Cost vectors by domain. Each spans approximately three orders
of magnitude, reflecting the heavy-tailed nature of error costs in
high-stakes applications. Calibration sources are detailed in
\autoref{sec:app-calibration}.}
\label{tab:costs}
\end{table}

The calibration of each cost level draws on multiple sources:
regulatory penalties (SEC SAB~99, SOX~\S906, FINRA, FDA, EU~AI~Act),
litigation data (securities class actions, medical malpractice
settlements), operational loss databases (ORX for banking, IBM/Ponemon
for data breaches), and published case studies of AI failures in
production \citep{aircanada2024,prosperspark2023}. Detailed
calibration evidence is provided in \autoref{sec:app-calibration}.

\paragraph{Severity assignment.}
We do not collect new human annotations. For each dataset, we
define deterministic rules that map an existing structured field
of the corpus (clause type, deal-point category, fixed
hypothesis, academic specialty, etc.) to a severity tier. The
tier reflects the cost of \emph{erring on that instance}, not a
property of the question: a CUAD extraction is labelled critical
when missing the clause exposes the buyer to material legal
risk, not because the clause text is verbose. The rule-based
approach is reproducible and removes per-instance annotator
subjectivity. We validate each rule system against an
independent LLM annotator (Claude Sonnet~4) using the severity
rubric as a one-shot prompt, reporting Cohen's $\kappa$ \citep{cohen1960} and
adjacent agreement
($\pm 1$ level).

% ----------------------------------------------------------------------------
\subsection{Models}
\label{sec:models}

We evaluate $12$ open-weight checkpoints chosen to span the
deployed LLM design space: $8$\,B to $120$\,B parameters, dense
and mixture-of-experts, base and reasoning-tuned, FP16, FP8 and
MXFP4 quantizations. Concretely: Granite-3.2-8B, Phi-4,
Gemma-3-12B/27B, Qwen3-14B/30B-A3B, QwQ-32B, GPT-OSS-20B/120B,
Qwen2.5-72B, Llama-3.3-70B and DeepSeek-R1-Distill-70B. All
inference runs on a $3\times$RTX~6000~Ada workstation via vLLM.
The exact checkpoint identifiers, quantization details, and the
inference hardware configuration are in
\autoref{sec:app-models}.

\subsection{Evaluation protocol}
\label{sec:protocol}

Each dataset is evaluated under an \emph{original} prompt
(reproducing the dataset's published protocol) and a
\emph{standard} HELM-style prompt \citep{liang2023helm} as an
ablation. Predictions are scored by a cascade (MCQ letter, yes/no,
numeric at $5\%$ tolerance, exact, fuzzy). For each (model,
dataset) pair we compute $\hat{p}{=}1{-}\mathrm{acc}$, the
severity profile $\hat{\bpi}$, and the actuarial measures
$\widehat{\ES}$, $\VaR_{0.95}$, $\TVaR_{0.95}$, $R$, $u^\ast$ via
$10^5$ Monte Carlo replications at $n{=}10{,}000$ outputs. Wilson
score intervals on accuracy, Goodman simultaneous intervals on
$\hat{\bpi}$, and percentile bootstrap intervals on the actuarial
measures are computed at $\alpha{=}0.95$ throughout. Full
methodological detail (scoring cascade, CI formulas, paired
bootstrap, per-domain aggregation) is in
\autoref{sec:app-protocol}; the cost-calibration robustness check
is in \autoref{sec:app-robustness}. The analysis is organised
around three research questions:

\begin{description}[leftmargin=0pt,labelindent=0pt]
  \item[RQ1 (Ranking divergence and inversions).] How closely aligned
    are model rankings by accuracy and by $\ES$ within a domain, and
    do concrete pairs invert? We report Kendall $\tau$
    \citep{kendall1938} between the
    two rank orders together with its $p$-value, plus the count and
    dollar magnitude of strict inversions where
    $\mathrm{acc}(A) > \mathrm{acc}(B)$ but $\ES(A) > \ES(B)$.
  \item[RQ2 (Severity vs.\ frequency variance).] How much of the
    inter-model variance in $\ES$ is attributable to the severity
    profile $\bpi$ versus the error rate $p$? We report a variance
    decomposition that holds one factor at the domain mean while
    letting the other vary.
  \item[RQ3 (Routing benefit).] Does the cost reduction from
    human-in-the-loop routing covary with the tail heaviness of the
    domain's cost distribution? We report the per-domain mean cost
    reduction and the Spearman correlation with the coefficient of
    variation of $\mathbf{c}$.
\end{description}

% ============================================================================
% 5. RESULTS
% ============================================================================

\section{Results}
\label{sec:results}

We report results on the $132$ (model, dataset) cells, each
evaluated on the \emph{full} test corpus. All risk measures are
computed with $n{=}10{,}000$ outputs per evaluation period,
$20{,}000$ Monte Carlo replications, and $\alpha{=}0.95$. The
complete per-pair table is provided in
\autoref{sec:app-full-results}. The headline numbers below
come from $12$ open-weight local models on $11$ benchmarks, with
per-cell sample sizes ranging from $150$ (FinanceBench) to
$182{,}822$ (MedMCQA).

\subsection{RQ1: Ranking alignment and dollar-magnitude inversions}
\label{sec:rq1}

\paragraph{Rankings agree, contrary to the small-sample claim.}
At full-corpus scale, accuracy and $\ES$ rankings are positively
correlated in every domain: per-domain Kendall $\tau$ is
$0.58$ in finance ($p{=}0.009$), $0.73$ in insurance
($p{=}5{\times}10^{-4}$), $0.97$ in legal ($p{=}5{\times}10^{-8}$),
and $0.94$ in medical ($p{=}3{\times}10^{-7}$)
(\autoref{tab:h1}). The dramatic divergence reported by
\citet{beauchemin2026position} on an $82$-query insurance corpus,
and the $\tau \approx 0.4$ we ourselves obtained on a
$1$\,K-per-dataset pilot run, do not reproduce at full corpus.
A subsample-from-full ablation
(\autoref{sec:app-subsample-ablation}) shows that the per-cell
sample size is the dominant driver of the $\tau$ shift in
finance and insurance: at $n{=}1000$ per cell the mean $\tau$
drops to $0.52$ and $0.45$ respectively, recovering the pilot
regime, while medical and legal remain at $\tau \geq 0.91$. The
prior divergence claim is therefore a small-sample artefact in
finance and insurance, not a property of the framework, and
does not arise in medical or legal at any tested sample
size.

\paragraph{Dollar magnitudes are still missing from accuracy.}
The framework's residual value is what an accuracy-only
leaderboard cannot show: the cost of the rank that is correct.
We detect $74$ strict accuracy-vs-$\ES$ inversions where
$\mathrm{acc}(A) > \mathrm{acc}(B)$ \emph{and} $\ES(A) > \ES(B)$,
out of $726$ ordered model pairs ($10.2\%$ of all pairs). The
gaps have a heavy right tail: median $\$1.9$\,M, $75$th
percentile $\$5.4$\,M, max $\$28.7$\,M, summing to $\$0.32$\,B
of mis-attributed exposure per $10{,}000$ outputs. The largest
is on JudgeBERT: DeepSeek-R1-70B edges Llama-3.3-70B by $+0.7$
accuracy points yet costs $\$28.7$\,M more; on ContractNLI,
Llama-3.3-70B is $+2.5$ over Granite-3.2-8B for $+\$21.9$\,M;
on TAT-QA, qwen2.5-72b is $+0.2$ over DeepSeek-R1-70B for
$+\$8.7$\,M. The most visible \emph{rank flip} is in finance:
DeepSeek-R1-70B is rank $4$ by accuracy and rank $1$ by $\ES$,
driven by a severity profile concentrated on negligible/minor
levels (\autoref{fig:ranking_finance}).

\paragraph{Domain-level aggregation reduces inversions to 13.}
The $74$ inversions are at the per-(model, dataset) level. When
ranks are aggregated to the domain (sample-weighted across the
domain's datasets) only $13$ inversions remain: $11$ in finance
(largest $\$7.5$\,M), $1$ in medical ($\$0.3$\,M), $1$ in
insurance ($\$0.8$\,M), and $0$ in legal. At the level on which
production deployments are scoped (a domain, not a single
dataset), accuracy and cost ranking agree to within a few model
positions. The framework's contribution is therefore not
overturning leaderboards: it attaches a defensible dollar
magnitude to the gap between two adjacent ranks. The takeaway
is no longer \emph{\enquote{accuracy gets the leaderboard
wrong}}; it is \emph{\enquote{accuracy doesn't tell you how much
being wrong costs}}.

\begin{table}[t]\centering
\begin{tabular}{lrrr}
\toprule
Domain & $\tau$ & $p$-value & 95\% bootstrap CI \\
\midrule
  Finance   & 0.58 & 0.009              & $[0.05,\,0.97]$ \\
  Insurance & 0.73 & $5{\times}10^{-4}$ & $[0.37,\,0.97]$ \\
  Legal     & 0.97 & $5{\times}10^{-8}$ & $[0.81,\,1.00]$ \\
  Medical   & 0.94 & $3{\times}10^{-7}$ & $[0.79,\,1.00]$ \\
\bottomrule
\end{tabular}
\caption{Kendall $\tau$ between accuracy and $\widehat{\ES}$
rankings per domain ($n_{\text{models}}{=}12$, full 11-corpus
evaluation, $10{,}000$ bootstrap resamples over models). The
point estimate is moderate for finance ($\tau{=}0.58$) and
strong for the other three domains ($\tau \geq 0.73$); the
bootstrap CIs are wide for finance and insurance, consistent
with the small number of ranked models.}
\label{tab:h1}
\end{table}

\begin{figure*}[t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
    yscale=0.62, xscale=1.0,
    every node/.style={font=\footnotesize},
    rise/.style={draw=sevNeg, line width=1.4pt},
    flat/.style={draw=black!30, line width=0.5pt},
    dot/.style={circle, fill=black!70, inner sep=1.1pt},
]
  \node[font=\footnotesize\bfseries, anchor=south west] at (-3.4,-0.1) {Accuracy};
  \node[font=\footnotesize\bfseries, anchor=south east] at (12.4,-0.1) {Expected loss $\widehat{\ES}$ per 10\,k outputs};
  % Sample-size-weighted finance ranks (n=2960 = 150+1147+1663 per model).
  \draw[flat] (0,-1) -- (9,-2);
  \draw[flat] (0,-2) -- (9,-4);
  \draw[flat] (0,-3) -- (9,-6);
  \draw[rise] (0,-4) -- (9,-1);
  \draw[flat] (0,-5) -- (9,-7);
  \draw[flat] (0,-6) -- (9,-5);
  \draw[flat] (0,-7) -- (9,-8);
  \draw[flat] (0,-8) -- (9,-9);
  \draw[rise] (0,-9) -- (9,-3);
  \draw[flat] (0,-10) -- (9,-10);
  \draw[flat] (0,-11) -- (9,-11);
  \draw[flat] (0,-12) -- (9,-12);
  % left labels: accuracy rank order, value before name
  \node[left] at (-0.2,-1)  {37.8\%\enskip Gemma-3-27B};
  \node[left] at (-0.2,-2)  {36.7\%\enskip Qwen3-14B};
  \node[left] at (-0.2,-3)  {36.4\%\enskip Qwen3-30B-A3B};
  \node[left] at (-0.2,-4)  {35.6\%\enskip \textcolor{sevNeg}{\bfseries DeepSeek-R1-70B}};
  \node[left] at (-0.2,-5)  {34.8\%\enskip Gemma-3-12B};
  \node[left] at (-0.2,-6)  {34.4\%\enskip Granite-3.2-8B};
  \node[left] at (-0.2,-7)  {34.2\%\enskip Llama-3.3-70B};
  \node[left] at (-0.2,-8)  {33.7\%\enskip QwQ-32B};
  \node[left] at (-0.2,-9)  {33.2\%\enskip \textcolor{sevNeg}{\bfseries GPT-OSS-20B}};
  \node[left] at (-0.2,-10) {32.9\%\enskip GPT-OSS-120B};
  \node[left] at (-0.2,-11) {31.6\%\enskip Qwen2.5-72B};
  \node[left] at (-0.2,-12) {29.6\%\enskip Phi-4};
  % right labels: E[S] rank order, name then value
  \node[right] at (9.2,-1)  {\textcolor{sevNeg}{\bfseries DeepSeek-R1-70B}\enskip \$25.1\,M};
  \node[right] at (9.2,-2)  {Gemma-3-27B\enskip \$30.2\,M};
  \node[right] at (9.2,-3)  {\textcolor{sevNeg}{\bfseries GPT-OSS-20B}\enskip \$32.1\,M};
  \node[right] at (9.2,-4)  {Qwen3-14B\enskip \$32.2\,M};
  \node[right] at (9.2,-5)  {Granite-3.2-8B\enskip \$32.3\,M};
  \node[right] at (9.2,-6)  {Qwen3-30B-A3B\enskip \$32.6\,M};
  \node[right] at (9.2,-7)  {Gemma-3-12B\enskip \$33.0\,M};
  \node[right] at (9.2,-8)  {Llama-3.3-70B\enskip \$33.3\,M};
  \node[right] at (9.2,-9)  {QwQ-32B\enskip \$33.3\,M};
  \node[right] at (9.2,-10) {GPT-OSS-120B\enskip \$33.5\,M};
  \node[right] at (9.2,-11) {Qwen2.5-72B\enskip \$34.1\,M};
  \node[right] at (9.2,-12) {Phi-4\enskip \$35.3\,M};
  \foreach \y in {1,...,12}{ \node[dot] at (0,-\y){}; \node[dot] at (9,-\y){}; }
\end{tikzpicture}%
}%
\caption{Finance: accuracy rank (left) vs.\ $\widehat{\ES}$ rank
(right) across 12 models, full corpus ($\tau{=}0.58$). Two models
rise sharply when ranked by dollar cost: DeepSeek-R1-70B
($4{\to}1$) and GPT-OSS-20B ($9{\to}3$).}
\label{fig:ranking_finance}
\end{figure*}

\subsection{RQ2: Severity vs.\ frequency variance}
\label{sec:rq2}

We decompose inter-model $\sigma^2_{\ES}$ into a severity-only
($\sigma^2_{\bpi}$) and a frequency-only ($\sigma^2_p$) component.
Across the four domains the severity share spans most of the
$[0, 1]$ interval: $0.81$ in finance, $0.37$ in legal, $0.06$ in
insurance, $0.05$ in medical (\autoref{fig:variance}).

\paragraph{Mechanism.}
The spread reflects the relative dispersion of the two inputs to
$\ES = n\, p\, \mu_X$. Across the $12$ models within a domain,
both the error rate $p$ and the mean severity
$\mu_X = \sum_k c_k \pi_k$ vary; the dominant component is the
one with the larger coefficient of variation. In finance, $p$ is
tightly clustered (CV $0.03$, range $0.62$--$0.70$) while $\mu_X$
moves twice as much (CV $0.07$); severity dominates. In medical
the relation reverses: $p$ has CV $0.74$ (range $0.01$ on DDI to
$0.36$ on HEAD-QA) while $\mu_X$ has CV $0.17$, so frequency
dominates by an order of magnitude. Insurance follows the
medical pattern ($p$-CV $0.67$, $\mu_X$-CV $0.17$); legal lies
between ($p$-CV $0.33$, $\mu_X$-CV $0.26$).

\paragraph{Practical implication.}
In finance, models differ mostly in \emph{which} errors they
make: identical accuracy can hide an order-of-magnitude dollar
exposure depending on the severity profile. In medical and
insurance, models differ almost entirely in \emph{how often}
they err: the severity profiles are nearly indistinguishable, so
error rate \emph{is} a faithful rank. The lever that most
reduces $\ES$ is therefore domain-specific: in finance, the
marginal dollar is spent improving severity selection (avoid the
worst errors); in medical and insurance, it is spent improving
accuracy outright. Legal sits between. The decomposition is
sensitive to the cost vector and to the model mix; we report it
as descriptive rather than confirmatory. Robustness checks:
$\pm 20\%$ perturbation of every cost level
(\autoref{sec:app-robustness}) does not change the cross-domain
ordering, and excluding the two saturated corpora (DDI 2013,
PrivacyQA) leaves $p$-CV and $\mu_X$-CV essentially unchanged
in finance, medical, and legal; only insurance reduces to a
single-corpus setting where $p$-CV drops from $0.67$ to $0.04$
because JudgeBERT alone shows little per-model error-rate
spread.

\subsection{RQ3: Routing tradeoff}
\label{sec:rq3}

The routing-as-reinsurance analysis turns the framework into an
operational decision tool: given a retention threshold $d{=}c_3$
(defer \enquote{major} and \enquote{critical} errors) and a human-review cost
$h{=}\$50$, how much $\ES$ is removed per fraction of queries
deferred? The answer is domain-stratified
(\autoref{tab:h4}, \autoref{fig:routing}). Finance is the
\enquote{cheap routing} regime: deferring the median $18\%$ of queries
removes $84.6\%$ of $\ES$, a leverage ratio of $4.7{:}1$. The
three other domains live in a \enquote{deep routing} regime, requiring
deferral of $41$--$53\%$ of the queries but extracting near-total
reductions ($92$--$98\%$). Insurance is the steepest combination
($98\%$ at $41\%$ routing), consistent with its heavy-tailed cost
vector (CV $= 1.63$) absorbing the deferred
tail. The pattern is robust to a $\pm 20\%$ multiplicative
perturbation of every cost level (\autoref{sec:app-robustness}):
the minimum Spearman correlation of per-model rankings under
perturbation is $0.90$ in finance and $1.00$ in the other three
domains.

\paragraph{Tail risk premium.}
The actuarial tail measures $\VaR_{0.95}$ and $\TVaR_{0.95}$
track the per-model $\ES$ ranking almost perfectly within every
domain (Kendall $\tau \geq 0.97$): they do not reshuffle models.
What they add is the size of the reserve a deployment should hold
above the expected loss. Median $\TVaR_{0.95}/\widehat{\ES}$
ratios are $1.08$ in finance, $1.06$ in legal, $1.13$ in medical
and $1.19$ in insurance, with the largest models in insurance
reaching $1.25$. In dollar terms, a $\$465$\,M-$\ES$ JudgeBERT
deployment carries an additional $\$15$--$\$20$\,M tail
contingency at the $95\%$ level. The insurance domain's heavier
ratio reflects its higher cost-vector CV.

\paragraph{Multi-period ruin reserve.}
For a multi-period deployment with a net-revenue safety loading
$\theta{=}0.2$ and a ruin tolerance $\psi^\ast{=}0.01$, the
Lundberg reserve $u^\ast = -\ln(\psi^\ast)/R$ gives the dollar
buffer that prevents bankruptcy with $99\%$ probability over an
infinite horizon. Median $u^\ast$ values per $10{,}000$ outputs
per period are $\$0.8$\,M (finance), $\$2.2$\,M (legal),
$\$2.9$\,M (insurance) and $\$6.2$\,M (medical). The medical
reserve is dominated by HEAD-QA, where the worst three models
score under $1\%$ accuracy and push $\mu_X$ above $\$170$\,k.
Reserves vary little across models within a domain (range $< 10\%$ of
the median in all four), confirming that ruin-style reserves
are driven primarily by the cost vector and the domain-wide
error distribution rather than by model choice.

\begin{table}[t]\centering
\begin{tabular}{lrrr}
\toprule
Domain & CV($c$) & $\rho_{1/2}$ & $\widehat{\ES}$ reduction \\
\midrule
  Finance   & 1.51 & 0.18 & 84.6\,\% \\
  Insurance & 1.63 & 0.41 & 98.3\,\% \\
  Legal     & 1.51 & 0.53 & 98.0\,\% \\
  Medical   & 1.51 & 0.51 & 92.0\,\% \\
\bottomrule
\end{tabular}
\caption{HITL routing tradeoff by domain ($n_{\text{models}}{=}12$,
retention threshold $d{=}c_3$, human review cost $h{=}\$50$).
$\rho_{1/2}$ is the median fraction of queries deferred. Finance
is the cheap-routing regime; the others are deep-routing.}
\label{tab:h4}
\end{table}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    font=\footnotesize,
    yscale=0.75,
    sevdot/.style={circle, fill=sevMaj, draw=sevMaj, inner sep=2.2pt},
    freqdot/.style={circle, fill=accentBlue, draw=accentBlue, inner sep=2.2pt},
    spine/.style={draw=black!18, line width=0.8pt, line cap=round},
]
  % horizontal spectrum spines, one per domain
  \foreach \y in {1,2,3,4} \draw[spine] (0,-\y) -- (6,-\y);
  % 50% reference tick (dotted)
  \draw[dotted, black!40] (3,-0.4) -- (3,-4.6);
  % axis tick labels at 0/50/100
  \node[font=\scriptsize, text=black!50] at (0,-4.8) {0\%};
  \node[font=\scriptsize, text=black!50] at (3,-4.8) {50\%};
  \node[font=\scriptsize, text=black!50] at (6,-4.8) {100\%};
  % domain labels (left) -- sorted by severity share descending
  \node[anchor=east] at (-0.15,-1) {Finance};
  \node[anchor=east] at (-0.15,-2) {Legal};
  \node[anchor=east] at (-0.15,-3) {Insurance};
  \node[anchor=east] at (-0.15,-4) {Medical};
  % markers: x = share * 0.06 (axis 0..6 for 0..100%)
  % Finance 81/19
  \node[sevdot]  at (4.86,-1) {};
  \node[freqdot] at (1.14,-1) {};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=sevMaj]    at (4.86,-1) {81\%};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=accentBlue] at (1.14,-1) {19\%};
  % Legal 37/63
  \node[sevdot]  at (2.22,-2) {};
  \node[freqdot] at (3.78,-2) {};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=sevMaj]    at (2.22,-2) {37\%};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=accentBlue] at (3.78,-2) {63\%};
  % Insurance 6/94
  \node[sevdot]  at (0.36,-3) {};
  \node[freqdot] at (5.64,-3) {};
  \node[anchor=south west, yshift=2pt, xshift=-2pt, font=\scriptsize, text=sevMaj] at (0.36,-3) {6\%};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=accentBlue] at (5.64,-3) {94\%};
  % Medical 5/95
  \node[sevdot]  at (0.30,-4) {};
  \node[freqdot] at (5.70,-4) {};
  \node[anchor=south west, yshift=2pt, xshift=-2pt, font=\scriptsize, text=sevMaj] at (0.30,-4) {5\%};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=accentBlue] at (5.70,-4) {95\%};
\end{tikzpicture}
\caption{Inter-model variance in $\widehat{\ES}$ split into a
\textcolor{sevMaj}{severity component $\sigma^2_{\bpi}/\sigma^2$}
and a \textcolor{accentBlue}{frequency component
$\sigma^2_p/\sigma^2$}. Each domain occupies its own spectrum;
the dots are the two complementary shares. Severity dominates in
finance; medical and insurance concentrate almost entirely on
frequency; legal sits between.}
\label{fig:variance}
\end{figure}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    font=\footnotesize,
    yscale=0.75,
    routedot/.style={circle, fill=black!55, draw=black!55, inner sep=2.2pt},
    reddot/.style={circle, fill=accentBlue, draw=accentBlue, inner sep=2.2pt},
    spine/.style={draw=black!18, line width=0.8pt, line cap=round},
    gap/.style={draw=accentBlue!40, line width=2.0pt, line cap=round},
]
  % spine 0..100
  \foreach \y in {1,2,3,4} \draw[spine] (0,-\y) -- (6,-\y);
  % column hint
  \node[font=\footnotesize, text=black!55, anchor=west] at (0,-0.15) {routed $\rho_{1/2}$};
  \node[font=\footnotesize, text=accentBlue, anchor=east] at (6,-0.15) {$\widehat{\ES}$ reduction};
  % domains (sorted by leverage = reduction - routed)
  \node[anchor=east] at (-0.15,-1) {Finance};
  \node[anchor=east] at (-0.15,-2) {Insurance};
  \node[anchor=east] at (-0.15,-3) {Legal};
  \node[anchor=east] at (-0.15,-4) {Medical};
  % Finance: 18 -> 85
  \draw[gap] (1.08,-1) -- (5.10,-1);
  \node[routedot] at (1.08,-1) {};
  \node[reddot]   at (5.10,-1) {};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=black!60]   at (1.08,-1) {18\%};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=accentBlue] at (5.10,-1) {85\%};
  % Insurance: 41 -> 98
  \draw[gap] (2.46,-2) -- (5.88,-2);
  \node[routedot] at (2.46,-2) {};
  \node[reddot]   at (5.88,-2) {};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=black!60]   at (2.46,-2) {41\%};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=accentBlue] at (5.88,-2) {98\%};
  % Legal: 53 -> 98
  \draw[gap] (3.18,-3) -- (5.88,-3);
  \node[routedot] at (3.18,-3) {};
  \node[reddot]   at (5.88,-3) {};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=black!60]   at (3.18,-3) {53\%};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=accentBlue] at (5.88,-3) {98\%};
  % Medical: 51 -> 92
  \draw[gap] (3.06,-4) -- (5.52,-4);
  \node[routedot] at (3.06,-4) {};
  \node[reddot]   at (5.52,-4) {};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=black!60]   at (3.06,-4) {51\%};
  \node[anchor=south, yshift=2pt, font=\scriptsize, text=accentBlue] at (5.52,-4) {92\%};
  % x-axis tick marks at 0/50/100
  \foreach \x/\v in {0/0, 3/50, 6/100} \node[font=\scriptsize, text=black!50] at (\x,-4.7) {\v\%};
\end{tikzpicture}
\caption{Dumbbell of \textcolor{black!55}{routed share
$\rho_{1/2}$} vs.\ \textcolor{accentBlue}{$\widehat{\ES}$
reduction} by domain ($d{=}c_3$, $h{=}\$50$, full corpus, 12
models). Finance shows the longest leverage gap: $18\%$ of queries
deferred removes $85\%$ of expected loss. The other three domains
need roughly half the queries deferred to extract near-total
cuts.}
\label{fig:routing}
\end{figure}

% ============================================================================
% 7. DISCUSSION
% ============================================================================

\section{Discussion}
\label{sec:discussion}

\subsection{The framework, not the model, is what we are evaluating}

The $12$ open-weight checkpoints we benchmark span $8$\,B to
$120$\,B parameters, dense and MoE, base and reasoning-tuned,
FP16 and FP8 and MXFP4. They are not interchangeable on
accuracy, $\ES$, or $\bpi$, yet the three framework outputs
surface the same kind of signal regardless of which model we
look at: a dollar-magnitude inversion in RQ1 that an accuracy
leaderboard cannot expose, a domain-specific severity-vs-
frequency variance share in RQ2, and a routing tradeoff in RQ3
that converts the framework's outputs into an operating point
(\enquote{cheap} or \enquote{deep} routing). The contribution is therefore
not \enquote{model $X$ is best} but \enquote{the framework reveals
decision-relevant cost structure that generalises across the
model design space we tested}. This is the empirical
complement to the conceptual demonstration of
\citet{beauchemin2026position}, which showed feasibility on a
single $82$-query insurance corpus with one model.

That said, one cross-model regularity is visible: the four
reasoning-tuned checkpoints (DeepSeek-R1-Distill-70B, QwQ-32B,
Qwen3-14B, Qwen3-30B-A3B) achieve lower mean $\widehat{\ES}$
than the eight non-reasoning models in every domain
($-\$112.8$\,M in medical, $-\$15.4$\,M in legal, $-\$6.5$\,M
in insurance, $-\$2.2$\,M in finance, per $10{,}000$ outputs).
The gap is largest where frequency dominates the variance
(medical, insurance): in those regimes, reasoning-tuned models
also have systematically lower $p$. This is descriptive rather
than causal: three of the four reasoning models are from
Alibaba (Qwen3-14B, Qwen3-30B-A3B, QwQ-32B) and Qwen2.5-72B
(same vendor, non-reasoning) lands among the more expensive
models in medical, so vendor and post-training procedure are
confounded with the reasoning label. The alignment between the
variance-share decomposition and the model-class gap is
consistent with the proposed mechanism but does not establish
it.

\subsection{What the actuarial framework adds over accuracy}

The compound loss model produces three families of quantities that
accuracy alone does not provide. The expected loss $\ES$ summarises
the per-period economic exposure under a given $\mathbf{c}$; the
tail measures $\VaR_{0.95}$ and $\TVaR_{0.95}$ characterise the
loss in adverse scenarios; the Lundberg reserve $u^\ast$ translates
a target ruin tolerance into a dollar budget. These outputs are
expressed in the same units that risk-management functions
(Basel~II/III, SOX, COSO) already use, which can lower the
translation cost between AI evaluation and enterprise risk
processes. The actuarial decomposition therefore complements
accuracy rather than replaces it: the error rate $p$ remains a
primary input, but $\bpi$ and $\mathbf{c}$ are needed to recover
the economic picture.

\subsection{When does severity matter most?}

The RQ2 mechanism (\autoref{sec:rq2}) has a direct
implication for procurement and model-selection workflows.
In \emph{severity-dominated} regimes (finance here, with
$\sigma^2_{\bpi}$-share $0.81$), two models with the same
accuracy can carry order-of-magnitude different dollar exposures:
ranking on accuracy alone leaves money on the table and
the model-selection committee must inspect $\bpi$. In
\emph{frequency-dominated} regimes (medical $0.05$, insurance
$0.06$), the $12$ models cluster on essentially the same
severity profile across all datasets in the domain, so the
error-rate ordering is already cost-faithful; the actuarial
overhead pays back only on tail-risk and reserve computations
(\autoref{sec:rq3}, tail-risk premium). Legal sits between
($0.37$), with a smaller but non-trivial severity contribution.
Whether a deployment is in the severity- or frequency-dominated
regime is a domain property given the cost vector, not a
property of the model; this can be checked before committing to
a more elaborate evaluation pipeline.

\subsection{Human-in-the-loop routing}

Modelling deferral as excess-of-loss reinsurance gives a closed-form
breakeven cost for human review, $h^\ast = \mu_X^{\text{route}}$,
the mean severity of the routed errors. In our snapshot, even a
$\$50$ human-review cost is well below this breakeven for medical
and insurance taxonomies, with estimated $\widehat{\ES}$
reductions of $92\%$ (medical) and $98\%$ (insurance) when
\enquote{major} and \enquote{critical} errors are routed ($d=c_3$). We caution that the per-error severity used in routing
is a function of $\bpi$ estimated from observed errors, not of an
oracle severity classifier; the realised benefit depends on how
accurately routing decisions are made at inference time.

\subsection{Limitations of cost calibration}

Cost vectors are approximate. Real costs depend on context,
downstream action, detection probability, and whether an error
propagates. Our levels represent expected per-instance costs
conditioned on an undetected error. The robustness analysis (\autoref{sec:app-robustness})
suggests that $\pm 20\%$ perturbations do not overturn rankings in
the snapshot we report, but practitioners should recalibrate
$\mathbf{c}$ to their organization's loss data when available.

\subsection{Comparison with MQM scoring}

The closest precedent in NLP is the MQM weighting scheme
\citep{lommel2014mqm}, with weights $\{1, 5, 25\}$ for minor,
major and critical errors. Applied to our severity labels (with
$0$ for negligible), MQM-weighted model rankings track our
$\widehat{\ES}$ rankings closely: per-domain Kendall $\tau$
between MQM and $\widehat{\ES}$ is $0.82$ (finance), $0.97$
(medical), $1.00$ (legal) and $0.97$ (insurance). MQM correctly
recovers most of the cost ordering when the weights are roughly
proportional to actual cost ratios. The framework differs from
MQM in three places where the difference is consequential rather
than ordinal: costs are calibrated against external evidence
(regulatory penalties, litigation data, operational loss
reports) so the absolute dollar magnitudes of the $74$
inversions are interpretable, not just their sign; tail measures
($\VaR$, $\TVaR$) and Lundberg reserves cannot be derived from a
weighted mean alone; and the routing analysis quantifies the
cost-benefit trade-off of human review, which MQM does not
address.

% ============================================================================
% 8. CONCLUSION
% ============================================================================

\section{Conclusion}
\label{sec:conclusion}

We applied the actuarial frequency-severity framework of
\citet{beauchemin2026position} to a multi-model, multi-domain LLM
evaluation. The framework decomposes operational risk into the error
rate and the severity profile, and yields three families of measures
(expected loss, tail risk, reserves) that complement accuracy.

Across $12$ open-weight models and four high-stakes domains
evaluated at full corpus, accuracy and $\ES$ rankings are
strongly correlated everywhere (Kendall $\tau$ from $0.58$ to
$0.97$): the dramatic divergence reported on a single $82$-query
insurance pilot is a small-sample artefact, not a stable empirical
regularity. The framework's residual value lies elsewhere. It
surfaces $74$ dollar-magnitude inversions in which the
higher-accuracy model is the more expensive one; it decomposes
$\sigma^2_{\ES}$ into severity ($0.05$--$0.81$) and frequency
components that vary by domain, revealing whether the lever for
reducing $\ES$ is \enquote{which errors} or \enquote{how often}; and it
yields a routing tradeoff that cuts $\ES$ by $85$\,\% (finance,
at $18$\,\% routing) up to $92$--$98$\,\% (medical, legal,
insurance, at $41$--$53$\,\% routing). The pattern holds across
the heterogeneous model sample, supporting the claim that the
framework, not any single model choice, quantifies the cost
structure that accuracy hides. Extending to additional severity
taxonomies (code/security, content moderation) is the natural
next step; the released \seveval{} library and
\texttt{run\_full\_pipeline.sh} script make that extension
reproducible. We release the code, severity
annotations, and analysis scripts to support adoption of
severity-aware evaluation as a complement to accuracy in
high-stakes deployments.

% ============================================================================
% LIMITATIONS
% ============================================================================

\section*{Limitations}
\label{sec:limitations}

\paragraph{Cost calibration.}
Our cost vectors are grounded in regulatory and market evidence
but the mapping from severity tiers to dollar amounts is
approximate; organisations within a domain face different cost
structures by size and regulatory exposure. The robustness
analysis (\autoref{sec:app-robustness}) shows rankings are
stable under $\pm 20\%$ perturbation, but practitioners should
recalibrate to their context.

\paragraph{Binary correctness.}
Each output is classified as correct or erroneous, discarding
partial correctness: \$$14.0$\,B instead of \$$14.2$\,B is
\enquote{wrong} but far less consequential than \$$11.6$\,B.
Continuous severity scoring is left to future work.

\paragraph{Severity-rule provenance.}
Severity is derived from a structured field of the source corpus
(clause type, deal-point category, fixed hypothesis, academic
specialty, OPP-115 category, DDI relation type). For the three
finance benchmarks the field does not exist natively and we use
a defensible answer-type$\times$metric-type matrix; an
independent LLM annotator validates with adjacent agreement
$94\%$, $\kappa{=}0.46$. Three benchmarks initially considered
(MedQA, MedCalc-Bench, RAG~Insurance) were dropped because their
keyword-based labels could not be defended at the
critical-vs-major boundary.

\paragraph{Saturated corpora.}
Two of the new corpora are near-saturation: DDI~2013 is solved
at $\ge 99.9\%$ by every model (relation-type classification with
no class imbalance), and PrivacyQA is solved at $\ge 83\%$ with
$\widehat{\ES}$ below $\$36$\,M for all models. These corpora
contribute little inter-model variance and serve mainly as
robustness checks that the per-domain ranking does not flip when
a quasi-trivial benchmark is added.

\paragraph{Inversion rate.}
The $74$ dollar-magnitude inversions arise from $726$ ordered
within-domain model pairs ($\binom{12}{2}$ pairs across $11$
datasets), so $10\%$ of all comparisons. The framework surfaces a
minority of mispriced pairs, not a leaderboard reshuffle.

\paragraph{Detection assumption.}
$\widehat{\ES}$, $\TVaR$ and the routing analysis all assume
errors are detected and routed at inference time. In production,
detection probability is below $1$; the realised cost reduction
from routing is therefore an upper bound. Calibrating an
oracle-free severity classifier is left to future work.

\paragraph{Pilot vs.\ full corpus is not a controlled ablation.}
The $\tau \approx 0.4$ we obtained on a $1$\,K-per-dataset pilot
also used the three later-dropped corpora and the broken
\texttt{mistral-small-3} checkpoint. The shift to $\tau \in
[0.6, 1.0]$ at full corpus therefore mixes saturation with the
cleaner roster; we cannot separate the two contributions without
a controlled study.

\paragraph{Scope.}
We evaluate English-language benchmarks across four domains.
The framework is domain-agnostic but the cost calibrations and
severity taxonomies do not transfer directly to other domains
(e.g., code generation, content moderation) without
domain-expert input. The evaluation is a single-point-in-time
snapshot; capabilities drift, and continuous monitoring via
\seveval{} is needed in production.

% ============================================================================
% ETHICS STATEMENT
% ============================================================================

\section*{Ethics Statement}
\label{sec:ethics}

This work evaluates existing language models on publicly available
benchmarks and does not involve human subjects research. The severity
annotations assign economic cost estimates to error categories but do
not make normative claims about the relative value of human outcomes
(e.g., we do not equate a medical error with a financial error based
on dollar amounts). The cost calibrations are derived from publicly
available regulatory and market data and are intended as illustrative
defaults for risk budgeting, not as authoritative valuations of human
welfare.

We release all code, annotations, and evaluation results to support
reproducibility. The \seveval{} library is released under an open-source
license. We note that the severity framework, like any evaluation
metric, can be used to inform but not replace human judgment in
deployment decisions.

% ============================================================================
% ACKNOWLEDGMENTS
% ============================================================================

% \section*{Acknowledgments}
%
% This work was supported by [funding sources]. We thank
% [names] for helpful discussions on actuarial modeling and
% [names] for feedback on annotation guidelines.

% ============================================================================
% BIBLIOGRAPHY
% ============================================================================

\bibliography{references}
% \bibliographystyle set by acl.sty

% ============================================================================
% APPENDIX
% ============================================================================

\appendix

\section{Cost Calibration Evidence}
\label{sec:app-calibration}

% Include from separate file
\input{severity-annotation-guidelines}

\section{The \seveval{} Library}
\label{sec:app-library}

We release \seveval{}, an open-source Python library implementing
the actuarial evaluation pipeline. The library follows three design
choices : (1)~a HuggingFace-style \texttt{evaluate.Metric} subclass
(\texttt{CompoundLossMetric}) so the risk measures plug into existing
evaluation pipelines alongside accuracy; (2)~minimal core
dependencies (NumPy, SciPy), with pandas and matplotlib loaded
lazily for DataFrame export and plotting; (3)~domain-agnostic
design, with built-in taxonomies for finance, medicine, law and
insurance plus a registration API for custom domains. Six modules
cover compound loss simulation, risk measures, ruin theory, routing
analysis, severity taxonomies, and sensitivity analysis.

\subsection{Core API}

The primary entry point is the \texttt{evaluate()} function:

\begin{small}
\begin{verbatim}
import severity_eval

report = severity_eval.evaluate(
    predictions=preds,
    references=refs,
    severity_annotations=labels,
    cost_levels=[100, 1_000, 10_000, 100_000],
    n_queries=10_000,
    n_sim=100_000,
    alpha=0.95,
)
\end{verbatim}
\end{small}

\noindent The returned \texttt{SeverityReport} object contains the
accuracy, error rate, severity profile $\hat{\bpi}$, expected loss,
VaR, TVaR, and the full vector of simulated losses. It supports
serialization to JSON, pandas DataFrames, and \LaTeX{} tables.

\subsection{Modules}

The actuarial pipeline is split across six analytical modules:

\begin{itemize}
  \item \textbf{compound\_loss}: Monte Carlo simulation of
    $S = \sum_{i=1}^{N} X_i$ under the binomial-categorical model.
  \item \textbf{risk\_measures}: Empirical VaR, TVaR, and bootstrap
    confidence intervals.
  \item \textbf{ruin}: Lundberg adjustment coefficient (via Brent's
    method), reserve requirements, and Monte Carlo ruin probability
    estimation.
  \item \textbf{routing}: Excess-of-loss routing analysis with
    configurable retention threshold and human review cost.
  \item \textbf{taxonomy}: Built-in severity taxonomies for finance,
    medicine, law and insurance, with a
    registration API for custom domains.
  \item \textbf{sensitivity}: Perturbation analysis of cost vectors
    (one coordinate at a time, $\pm$ directions) to assess robustness
    of model rankings.
\end{itemize}

Three auxiliary modules support the pipeline: \texttt{api} (the
\texttt{evaluate()} entry point and the \texttt{SeverityReport}
dataclass), \texttt{metric} (the HuggingFace \texttt{evaluate.Metric}
wrapper) and \texttt{validation} (input checks for severity profiles).
A \texttt{visualization} module provides the matplotlib plotting
helpers used by the figures in this paper.

\subsection{HuggingFace Integration}

The module \texttt{severity\_eval.metric} exposes
\texttt{compute\_severity\_metrics()}, a flat-dict wrapper around
the core \texttt{evaluate()} call, and \texttt{CompoundLossMetric},
a subclass of \texttt{evaluate.Metric} that can be instantiated
directly. Both paths return the same payload (accuracy, error rate,
severity profile, expected loss, VaR, TVaR), so the risk measures
can be reported alongside accuracy in a standard evaluation
pipeline. The metric is not published on the HuggingFace Hub as a
loadable module in this release.

\subsection{Reproducing the experiments}

The \texttt{experiments/run\_full\_pipeline.sh} script orchestrates
the full evaluation matrix: it evaluates the open-weight models
locally via vLLM with FP8 (or MXFP4-native for the gpt-oss family)
checkpoints, runs the analysis
(\texttt{experiments/analysis.py}) and regenerates all paper
figures and tables. Stages are idempotent and can be resumed with
the corresponding \texttt{--skip-*} flag.

\section{Severity Annotation and Per-Domain Scales}
\label{sec:app-severity-by-dataset}

\paragraph{Inter-annotator agreement.}
On the FinanceBench rule-based annotation, Cohen's
$\kappa = 0.46$ with $65.30\%$ exact agreement and $94.00\%$
adjacent agreement against an independent LLM annotator ($150$
instances). The \texttt{experiments/validate\_severity\_llm.py}
script reproduces this measurement and can be applied to
additional datasets to extend the table.

\paragraph{Per-dataset mapping rules.}
We detail the deterministic mapping rules and one representative
instance per dataset. Each rule maps a \emph{structured} field
of the source corpus to a four-level tier; mappings never
re-interpret raw question text.

\paragraph{FinanceBench.}
Severity comes from cross-tabulating \emph{answer type} (dollar
amount in millions/billions, percentage, ratio, yes/no,
free-text) with \emph{metric type} (core financial statement,
operational, derived, descriptive). The gold answer to \enquote{Does
Corning have positive working capital based on FY2022 data?}
is \$$831$M positive; mis-stating the sign of a yes/no over a
core operational metric can flip a credit decision
(\emph{critical}). Mis-reporting Amazon's FY2017 DPO of
$93.86$ (a derived ratio over a non-core operational metric)
typically yields a smaller business consequence (\emph{minor}).

\paragraph{FinQA and TAT-QA.}
Same two-axis matrix as FinanceBench. Mis-answering \enquote{by what
percentage did average borrowings decrease from $2016$ to
$2017$?} (gold $-8.40\%$, FinQA) is a small-magnitude reasoning
slip on a derived ratio (\emph{minor}); mis-answering \enquote{What
does the deferred income taxes reflect?} (free-form span,
TAT-QA) yields a descriptive paraphrase error that does not
change a downstream accounting decision (\emph{negligible}).

\paragraph{HEAD-QA.}
Severity follows the dataset's six \emph{academic-category}
labels by clinical-decision impact: pharmacology
$\to$ critical, medicine and nursing $\to$ major, psychology
$\to$ minor, biology and chemistry $\to$ negligible. A wrong
answer on a pharmacology stem defining \enquote{Parameter} propagates
into every downstream biostatistics decision
(\emph{critical}); a wrong answer on a biology fact-recall
stem (e.g.\ \enquote{does carotid body stimulation decrease arterial
PCO\textsubscript{2}?}) carries little clinical-decision
exposure (\emph{negligible}).

\paragraph{CUAD.}
The dataset's $41$ clause types map to severity by contractual
risk, following due-diligence practice
\citep{guha2023legalbench}: indemnification and liability
limitation $\to$ critical, non-compete and IP assignment
$\to$ major, renewal and notice $\to$ minor, definitional
$\to$ negligible. Failing to extract a \enquote{Termination for
Convenience} clause exposes the buyer to unilateral
cancellation (\emph{critical}); failing on an \enquote{Audit Rights}
clause may delay compliance review but is rarely catastrophic
(\emph{major}).

\paragraph{MAUD and ContractNLI.}
Both reuse the CUAD clause-type risk schedule. An incorrect MAUD
answer on whether the target's securities are subject to a
\enquote{disproportionate impact} modifier in the Material Adverse
Effect definition directly mis-states a deal closing condition
(\emph{critical}). An incorrect ContractNLI label on the
hypothesis \enquote{Receiving Party may independently develop
information similar to Confidential Information} generally
turns one boilerplate provision into another, so the business
exposure of the error is small (\emph{negligible}).

\paragraph{JudgeBERT.}
Severity is the cross product of the simplified clause's legal
category (exclusion, indemnity, definition, \dots) and the
human-evaluator score on meaning preservation: clauses bearing
material legal weight (exclusions, consequences of
non-compliance) at low evaluator score $\to$ critical;
definitional clauses at high score $\to$ negligible.

\paragraph{MedMCQA, DDI~2013, PrivacyQA.}
\textbf{MedMCQA} reuses HEAD-QA's specialty mapping on the
\texttt{subject\_name} field. \textbf{DDI~2013} maps the native
relation type to severity directly: \emph{effect} $\to$
critical, \emph{advice}/\emph{mechanism} $\to$ major,
\emph{int} $\to$ minor, \emph{false} (no interaction) $\to$
negligible. \textbf{PrivacyQA} maps the question to an OPP-115
category \citep{wilson2016opp115} whose severity tier reflects
the corresponding regulatory-fine band (Data Security and
Third-Party Sharing $\to$ critical/major; first-party
collection, user choice, access/edit $\to$ minor; descriptive
sections $\to$ negligible).

% Per-domain cost vectors (Medical / Legal / Code / Moderation)
\input{severity-scales-all-domains}

\section{Open-Weight Models and Inference}
\label{sec:app-models}

\begin{table}[t]
\centering
\setlength{\tabcolsep}{4pt}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}llll@{}}
\toprule
Model & Params & Provider & Type \\
\midrule
Granite-3.2-8B          & 8\,B   & IBM       & General (FP16)  \\
GPT-OSS-20B             & 20\,B  & OpenAI    & General (MXFP4) \\
Phi-4                   & 14\,B  & Microsoft & General         \\
Gemma-3-12B             & 12\,B  & Google    & General         \\
Qwen3-14B               & 14\,B  & Alibaba   & Reasoning       \\
Gemma-3-27B             & 27\,B  & Google    & General         \\
Qwen3-30B-A3B           & 30\,B  & Alibaba   & MoE             \\
QwQ-32B                 & 32\,B  & Alibaba   & Reasoning       \\
DeepSeek-R1-Distill-70B & 70\,B  & DeepSeek  & Reasoning       \\
Llama-3.3-70B           & 70\,B  & Meta      & General         \\
Qwen2.5-72B             & 72\,B  & Alibaba   & General         \\
GPT-OSS-120B            & 120\,B & OpenAI    & General (MXFP4) \\
\bottomrule
\end{tabular}%
}%
\caption{The $12$ open-weight checkpoints evaluated. All are served
by vLLM (RedHatAI FP8, Qwen-native FP8, MXFP4 for gpt-oss, FP16 for
Granite). $70$--$72$\,B and $120$\,B models use TP$=2$; others
single-GPU.}
\label{tab:models}
\end{table}

\paragraph{Inference hardware and quantization.}
Local open-weight models are served by vLLM with FP8 quantization
(RedHatAI compressed-tensors packaging where available, Qwen-native
FP8 otherwise) and MXFP4 for the gpt-oss family. The
70-72\,B Llama, DeepSeek, and Qwen2.5 checkpoints and gpt-oss-120b
run with \texttt{tensor\_parallel\_size}=2 on two NVIDIA RTX 6000 Ada
48\,GB cards; the remaining open-weight models fit on a single card.
The granite-3.2-8b checkpoint is FP16 (no FP8 variant published). We
use adaptive inference parameters per dataset, with
\texttt{max\_new\_tokens} ranging from 16 (ContractNLI, binary
answers) to 256 (CUAD, clause extraction). Each model's vLLM
submission is chunked at 100 prompts per chunk so the per-dataset
output JSON is checkpointed and the run can be resumed or extended
from a partial file.

\section{Evaluation Protocol Details}
\label{sec:app-protocol}

This section expands on the body's \autoref{sec:protocol};
prompt templates per dataset are in \autoref{sec:app-prompts}.

\paragraph{Scoring cascade.}
Predictions are scored by the following cascade: option-letter
match for MCQ; yes/no extraction; numeric comparison at $5\%$
relative tolerance; exact text match after normalisation;
substring containment (min.\ $4$ chars); word-set overlap
(min.\ $3$ reference words).

\paragraph{Risk measures.}
For each (model, dataset) pair we compute
$\hat{p}{=}1{-}\mathrm{accuracy}$, the severity profile
$\hat{\bpi}$ from the empirical error distribution,
$\widehat{\ES}$ via \autoref{eq:expected_loss}, and the full
distribution of $S$ from $10^5$ Monte Carlo replications at
$n{=}10{,}000$ outputs per evaluation period (which yields
$\VaR_{0.95}$ and $\TVaR_{0.95}$). The Lundberg coefficient $R$
and reserve $u^\ast$ are computed at ruin tolerance
$\psi^\ast{=}0.01$; routing cost reduction is reported at
thresholds $d \in \{c_2, c_3\}$ with human review cost
$h{=}\$50$. All computations use the \seveval{} library
(\autoref{sec:app-library}).

\paragraph{Confidence intervals.}
Our benchmarks span three orders of magnitude in size
(financebench $n{=}150$, judgebert $n{=}1{,}483$, headqa
$n{=}12{,}751$, maud $n{=}39{,}231$), so per-dataset confidence
intervals on every reported quantity are essential to avoid
over-reading small-sample noise. For each (model, dataset) pair
we report:
\begin{itemize}
  \item Accuracy $\hat{p}$: Wilson score $95\%$ interval
    \citep{wilson1927ci}, which is well-behaved near
    $\hat{p}{=}0$ and $\hat{p}{=}1$ unlike the normal
    approximation.
  \item Severity profile $\hat{\bpi}$: Goodman simultaneous
    multinomial $95\%$ intervals on each $\hat{\pi}_k$, with the
    family-wise rate controlled via Bonferroni across $K{=}4$
    categories.
  \item $\widehat{\ES}$, $\VaR_{0.95}$, $\TVaR_{0.95}$:
    non-parametric percentile bootstrap with $B{=}1{,}000$
    resamples over the $10^5$-replication Monte Carlo loss
    vector, exposed through
    \texttt{severity\_eval.bootstrap\_ci}.
\end{itemize}

\paragraph{Pairwise comparisons and per-domain aggregation.}
For pairwise model comparisons within a dataset we use a paired
bootstrap on the per-instance loss vector
$L_i = \mathbb{1}\{\text{wrong}_i\} \cdot c(\text{severity}_i)$,
which controls for shared difficulty across instances. The
two-sample $z$-test on accuracy is reported only as a quick
sanity check; bootstrap intervals are the primary inferential
tool because they remain valid under the heavy-tailed cost
vector. Per-domain aggregates weight datasets by sample size to
avoid letting a small benchmark (financebench, $n{=}150$)
dominate a large one (maud, $n{=}39{,}231$) when the per-dataset
CIs differ by an order of magnitude.

\section{Prompt Templates}
\label{sec:app-prompts}

% Include from separate file
\input{prompt_documentation}

\section{Robustness to Cost Calibration}
\label{sec:app-robustness}

\paragraph{Single-coordinate $\pm 20\%$ perturbations.}
Perturbing each $c_k$ by $\pm 20\%$ one coordinate at a time
barely moves the per-domain $\ES$ rankings: the minimum Spearman
correlation between the base and perturbed rankings is $0.93$ in
finance and $\geq 0.99$ in legal, insurance, and medical (mean
$\geq 0.97$ in all four). Moderate uncertainty in $\mathbf{c}$
does not overturn rankings dominated by large differences in
$\bpi$ and $p$. The full sensitivity table is generated by
\texttt{experiments/analysis.py} via the \texttt{--robustness}
flag.

\paragraph{Spacing ratio sensitivity.}
The default cost vectors use a $10$x geometric spacing
($c_{k+1}/c_k = 10$, except insurance which uses $20, 5, 25$).
We re-rank models under three alternative spacings: $5$x
geometric, $20$x geometric, and a log-linear spacing
($1, 4, 15, 60$ relative to $c_1$). Spearman correlations
between each alternative and the default ranking are:

\begin{center}
\small
\begin{tabular}{lrrr}
\toprule
Domain    & $5$x & $20$x & log-linear \\
\midrule
Finance   & $0.92$ & $0.99$ & $0.85$ \\
Medical   & $0.99$ & $1.00$ & $0.99$ \\
Legal     & $1.00$ & $1.00$ & $1.00$ \\
Insurance & $0.99$ & $1.00$ & $0.99$ \\
\bottomrule
\end{tabular}
\end{center}

\noindent The log-linear spacing reshuffles finance the most
($\rho{=}0.85$), consistent with the high severity-share in that
domain. The medical, legal, and insurance rankings are stable
across all tested spacings because frequency dominates inter-model
variance and is unaffected by the cost-level ratio.

\paragraph{Lundberg sensitivity grid.}
The Lundberg reserve $u^\ast$ depends on the safety loading
$\theta$ (a conventional choice in actuarial reserving;
\citealp{klugman2012loss,dickson2016insurance}) and the ruin
tolerance $\psi^\ast$. Median per-domain $u^\ast$ values
($10^{6}$\,USD per $10{,}000$ outputs per period) across a
$3 \times 3$ grid:

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{l ccc ccc ccc}
\toprule
& \multicolumn{3}{c}{Finance} & \multicolumn{3}{c}{Medical} & \multicolumn{3}{c}{Insurance} \\
\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
$\theta \backslash \psi^\ast$ & $.005$ & $.01$ & $.05$
                              & $.005$ & $.01$ & $.05$
                              & $.005$ & $.01$ & $.05$ \\
\midrule
$0.1$ & $1.6$ & $1.4$ & $0.9$ & $13.3$ & $11.6$ & $7.5$ & $6.2$ & $5.4$ & $3.5$ \\
$0.2$ & $0.9$ & $0.8$ & $0.5$ & $7.1$  & $6.2$  & $4.0$ & $3.3$ & $2.9$ & $1.9$ \\
$0.5$ & ---   & ---   & ---   & $3.3$  & $2.9$  & $1.9$ & $1.6$ & $1.4$ & $0.9$ \\
\bottomrule
\end{tabular}%
}
\end{center}

\noindent Reserves scale as expected with both $\theta$ and
$\psi^\ast$. At $\theta = 0.5$ Lundberg's adjustment coefficient
$R$ has no positive root in finance (the safety loading exceeds
the achievable margin), so $u^\ast$ is undefined; reserves
should instead be taken from the Monte Carlo tail of $S$.
Reported values in the body use the actuarial defaults
$\theta = 0.2$ and $\psi^\ast = 0.01$.

\paragraph{Aggregation method sensitivity.}
Kendall $\tau$ values in \autoref{tab:h1} use an
\emph{unweighted} mean of per-(model, dataset) accuracy and
expected loss within each domain. With a \emph{sample-weighted}
mean (each dataset weighted by its corpus size), finance
$\tau$ rises from $0.58$ to $0.67$; medical and insurance
become $0.97$ each; legal is unchanged at $1.00$. The
slopegraph in \autoref{fig:ranking_finance} uses
sample-weighted aggregates; the table reports unweighted values
because they match the analysis script's
$\texttt{groupby("model").mean()}$ default. Both aggregations
agree on the qualitative claim that finance is the loosest
domain.

\section{Subsample Ablation}
\label{sec:app-subsample-ablation}

To separate the effect of per-cell sample size from the effect of
roster change (the pilot used $3$ later-dropped corpora and a
broken $\texttt{mistral-small-3}$ checkpoint), we subsample the
$132$ full-corpus prediction files uniformly without replacement
to $n_{\text{sub}}$ predictions per (model, dataset) cell, with
the same $12$ models and $11$ datasets in scope, and recompute
the per-domain Kendall $\tau$. The procedure is repeated over
$10$ random seeds.

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lrrrr}
\toprule
$n_{\text{sub}}$ & Finance & Medical & Legal & Insurance \\
\midrule
$200$  & $0.13 \pm 0.21$ & $0.84 \pm 0.05$ & $0.76 \pm 0.11$ & $0.11 \pm 0.19$ \\
$500$  & $0.36 \pm 0.15$ & $0.90 \pm 0.03$ & $0.86 \pm 0.06$ & $0.32 \pm 0.20$ \\
$1000$ & $0.52 \pm 0.08$ & $0.91 \pm 0.02$ & $0.92 \pm 0.03$ & $0.45 \pm 0.19$ \\
full   & $0.58$          & $0.94$          & $0.97$          & $0.73$          \\
\bottomrule
\end{tabular}%
}
\end{center}

\noindent At $n_{\text{sub}} = 1000$ (close to the original
$1$\,K-per-dataset pilot) the per-domain $\tau$ reproduces the
pilot $\tau \approx 0.4$ range in finance and insurance
($0.52$ and $0.45$ mean across $10$ seeds), while medical and
legal remain in the $\tau \geq 0.91$ regime. At
$n_{\text{sub}} = 200$ finance and insurance approach zero
correlation. The pilot divergence is therefore primarily a
finite-sample effect in the two domains where the framework
shows the most variance (finance: severity-driven, insurance:
small model spread), not a property of the framework or of the
specific datasets dropped.

\section{Full Results Tables}
\label{sec:app-full-results}

We split the per-(model, dataset) results across four
domain-specific tables (\autoref{tab:results-finance},
\autoref{tab:results-legal}, \autoref{tab:results-medical},
\autoref{tab:results-insurance}). Costs are in millions of US dollars
for $n{=}10{,}000$ queries. The full 95\% bootstrap CI columns and
the Wilson CI on accuracy are released in
\texttt{results/metrics.csv}. Rows are sorted by $\widehat{\ES}$
within each dataset (lower is better). Generated by
\texttt{experiments/analysis.py}.

\begin{table*}[t]\centering\caption{\textbf{Finance} per-(dataset, model) accuracy and risk
measures. $\widehat{\ES}$ and tail measures in \$M.}\label{tab:results-finance}
\begin{tabular}{@{}llrrrrr@{}}
\toprule
Dataset & Model & Acc.\ (\%) & $\hat{p}$ & $\widehat{\ES}$ & VaR$_{95}$ & TVaR$_{95}$ \\
\midrule
\multirow{12}{*}{\rotatebox{90}{financebench}}
  & qwen2.5-72b              & 50.7 & 0.493 & 22.2 & 24.1 & 24.7 \\
  & qwen3-14b                & 56.0 & 0.440 & 22.9 & 24.8 & 25.3 \\
  & llama-3.3-70b            & 48.7 & 0.513 & 23.0 & 24.9 & 25.4 \\
  & qwen3-30b-a3b            & 55.3 & 0.447 & 24.2 & 26.1 & 26.6 \\
  & qwq-32b                  & 47.3 & 0.527 & 26.7 & 28.6 & 29.2 \\
  & gemma-3-27b              & 40.7 & 0.593 & 26.8 & 28.7 & 29.2 \\
  & phi-4                    & 37.3 & 0.627 & 27.0 & 28.9 & 29.4 \\
  & gemma-3-12b              & 34.0 & 0.660 & 28.0 & 29.9 & 30.4 \\
  & deepseek-r1-distill-70b  & 26.0 & 0.740 & 30.6 & 32.0 & 32.4 \\
  & granite-3.2-8b           & 35.3 & 0.647 & 34.4 & 36.8 & 37.4 \\
  & gpt-oss-120b             & 22.7 & 0.773 & 37.0 & 39.0 & 39.5 \\
  & gpt-oss-20b              & 19.3 & 0.807 & 39.7 & 41.6 & 42.1 \\
\midrule
\multirow{12}{*}{\rotatebox{90}{finqa}}
  & deepseek-r1-distill-70b  & 23.4 & 0.766 & 38.6 & 40.7 & 41.2 \\
  & granite-3.2-8b           & 13.1 & 0.869 & 46.9 & 49.3 & 49.8 \\
  & qwen3-14b                & 13.5 & 0.865 & 47.4 & 49.6 & 50.2 \\
  & gemma-3-27b              & 11.1 & 0.889 & 47.4 & 49.6 & 50.2 \\
  & gemma-3-12b              & 10.4 & 0.896 & 48.6 & 50.8 & 51.4 \\
  & qwq-32b                  & 10.5 & 0.895 & 48.6 & 50.9 & 51.5 \\
  & qwen3-30b-a3b            & 11.6 & 0.884 & 49.0 & 51.3 & 51.9 \\
  & gpt-oss-20b              & 11.2 & 0.888 & 49.1 & 51.5 & 52.1 \\
  & gpt-oss-120b             & 10.1 & 0.899 & 49.4 & 51.8 & 52.4 \\
  & qwen2.5-72b              &  9.5 & 0.905 & 50.1 & 52.5 & 53.1 \\
  & llama-3.3-70b            &  9.9 & 0.901 & 50.2 & 52.5 & 53.2 \\
  & phi-4                    &  8.5 & 0.915 & 51.0 & 53.3 & 53.9 \\
\midrule
\multirow{12}{*}{\rotatebox{90}{tatqa}}
  & deepseek-r1-distill-70b  & 44.9 & 0.551 & 15.4 & 16.9 & 17.3 \\
  & gemma-3-27b              & 55.9 & 0.441 & 18.6 & 20.3 & 20.8 \\
  & gpt-oss-20b              & 49.6 & 0.504 & 19.7 & 21.4 & 21.9 \\
  & qwen3-30b-a3b            & 51.8 & 0.482 & 22.0 & 23.9 & 24.4 \\
  & granite-3.2-8b           & 48.9 & 0.511 & 22.1 & 24.0 & 24.5 \\
  & gpt-oss-120b             & 49.5 & 0.505 & 22.1 & 24.0 & 24.5 \\
  & llama-3.3-70b            & 49.6 & 0.504 & 22.5 & 24.4 & 24.9 \\
  & qwen3-14b                & 51.0 & 0.490 & 22.5 & 24.4 & 24.9 \\
  & gemma-3-12b              & 51.7 & 0.483 & 22.7 & 24.7 & 25.2 \\
  & qwq-32b                  & 48.5 & 0.515 & 23.4 & 25.3 & 25.8 \\
  & qwen2.5-72b              & 45.2 & 0.548 & 24.1 & 26.1 & 26.6 \\
  & phi-4                    & 43.5 & 0.565 & 25.3 & 27.3 & 27.9 \\
\bottomrule
\end{tabular}
\end{table*}

\begin{table*}[t]\centering\caption{\textbf{Legal} per-(dataset, model) accuracy and risk
measures. $\widehat{\ES}$ and tail measures in \$M.}\label{tab:results-legal}
\begin{tabular}{@{}llrrrrr@{}}
\toprule
Dataset & Model & Acc.\ (\%) & $\hat{p}$ & $\widehat{\ES}$ & VaR$_{95}$ & TVaR$_{95}$ \\
\midrule
\multirow{12}{*}{\rotatebox{90}{contractnli}}
  & qwen3-30b-a3b            & 84.5 & 0.155 &  53.3 &  58.2 &  59.5 \\
  & gpt-oss-20b              & 83.8 & 0.162 &  68.5 &  74.1 &  75.5 \\
  & gpt-oss-120b             & 84.3 & 0.157 &  70.7 &  76.5 &  78.0 \\
  & qwq-32b                  & 81.7 & 0.183 &  73.4 &  79.2 &  80.7 \\
  & qwen3-14b                & 82.0 & 0.180 &  80.9 &  87.2 &  88.7 \\
  & qwen2.5-72b              & 82.3 & 0.177 &  91.5 &  98.1 &  99.8 \\
  & gemma-3-27b              & 82.7 & 0.173 &  92.3 &  98.9 & 100.6 \\
  & deepseek-r1-distill-70b  & 81.3 & 0.187 &  98.2 & 105.0 & 106.8 \\
  & gemma-3-12b              & 80.6 & 0.194 & 104.2 & 111.1 & 112.9 \\
  & granite-3.2-8b           & 72.9 & 0.271 & 122.5 & 130.0 & 132.0 \\
  & phi-4                    & 77.5 & 0.225 & 128.3 & 136.1 & 138.1 \\
  & llama-3.3-70b            & 75.5 & 0.245 & 144.4 & 152.5 & 154.5 \\
\midrule
\multirow{12}{*}{\rotatebox{90}{cuad}}
  & qwen2.5-72b              & 17.2 & 0.828 & 211.2 & 219.0 & 221.0 \\
  & gemma-3-12b              & 16.7 & 0.833 & 211.4 & 219.3 & 221.4 \\
  & gemma-3-27b              & 17.2 & 0.828 & 211.9 & 219.9 & 221.9 \\
  & llama-3.3-70b            & 15.5 & 0.845 & 212.2 & 220.2 & 222.2 \\
  & qwq-32b                  & 17.0 & 0.830 & 213.4 & 221.4 & 223.6 \\
  & gpt-oss-120b             & 18.1 & 0.819 & 213.6 & 221.5 & 223.6 \\
  & qwen3-30b-a3b            & 15.7 & 0.843 & 214.0 & 222.0 & 224.0 \\
  & gpt-oss-20b              & 17.1 & 0.829 & 214.5 & 222.6 & 224.7 \\
  & qwen3-14b                & 14.5 & 0.855 & 217.3 & 225.2 & 227.4 \\
  & phi-4                    & 12.6 & 0.874 & 218.6 & 226.7 & 228.8 \\
  & granite-3.2-8b           & 12.2 & 0.878 & 223.8 & 231.9 & 234.1 \\
  & deepseek-r1-distill-70b  &  4.3 & 0.957 & 232.1 & 240.2 & 242.4 \\
\midrule
\multirow{12}{*}{\rotatebox{90}{maud}}
  & gpt-oss-20b              & 82.6 & 0.174 &  34.6 &  38.2 &  39.1 \\
  & gpt-oss-120b             & 77.6 & 0.224 &  53.6 &  57.8 &  58.9 \\
  & qwen3-30b-a3b            & 64.9 & 0.351 & 117.8 & 124.6 & 126.3 \\
  & qwq-32b                  & 64.9 & 0.351 & 119.9 & 126.7 & 128.5 \\
  & qwen3-14b                & 64.7 & 0.353 & 125.5 & 132.5 & 134.3 \\
  & gemma-3-27b              & 54.8 & 0.452 & 211.3 & 220.5 & 222.9 \\
  & llama-3.3-70b            & 53.6 & 0.464 & 230.3 & 239.9 & 242.3 \\
  & gemma-3-12b              & 52.8 & 0.472 & 233.6 & 243.4 & 245.7 \\
  & granite-3.2-8b           & 49.1 & 0.509 & 274.9 & 285.4 & 288.1 \\
  & qwen2.5-72b              & 50.3 & 0.497 & 280.2 & 291.0 & 293.7 \\
  & deepseek-r1-distill-70b  & 19.1 & 0.809 & 499.8 & 513.0 & 516.4 \\
  & phi-4                    & 16.8 & 0.832 & 525.7 & 538.9 & 542.4 \\
\bottomrule
\end{tabular}
\end{table*}

\begin{table*}[t]\centering\caption{\textbf{Medical} per-(dataset, model) accuracy
and risk measures. $\widehat{\ES}$ and tail measures in \$M. DDI is
saturated by every model ($\widehat{\ES}{\le}\$1\,M$).}\label{tab:results-medical}
\begin{tabular}{@{}llrrrrr@{}}
\toprule
Dataset & Model & Acc.\ (\%) & $\hat{p}$ & $\widehat{\ES}$ & VaR$_{95}$ & TVaR$_{95}$ \\
\midrule
\multirow{12}{*}{\rotatebox{90}{headqa}}
  & gpt-oss-20b              & 86.0 & 0.140 &  145.3 &  158.0 &  161.2 \\
  & phi-4                    & 80.1 & 0.199 &  172.3 &  185.5 &  189.1 \\
  & gpt-oss-120b             & 75.1 & 0.249 &  257.8 &  274.4 &  278.6 \\
  & qwen3-30b-a3b            & 47.7 & 0.523 &  518.2 &  541.2 &  546.9 \\
  & qwq-32b                  & 36.5 & 0.635 &  627.6 &  652.3 &  658.4 \\
  & granite-3.2-8b           & 29.5 & 0.705 &  733.7 &  760.2 &  767.0 \\
  & qwen2.5-72b              & 25.8 & 0.742 &  752.2 &  778.8 &  785.3 \\
  & qwen3-14b                & 20.1 & 0.799 &  811.5 &  839.0 &  845.8 \\
  & deepseek-r1-distill-70b  & 15.0 & 0.850 &  833.4 &  860.9 &  867.8 \\
  & llama-3.3-70b            &  0.9 & 0.991 & 1006.7 & 1036.3 & 1044.1 \\
  & gemma-3-12b              &  0.4 & 0.996 & 1013.1 & 1042.5 & 1050.4 \\
  & gemma-3-27b              &  0.2 & 0.998 & 1016.3 & 1045.8 & 1053.4 \\
\midrule
\multirow{12}{*}{\rotatebox{90}{medmcqa}}
  & gpt-oss-20b              & 100.0 & 0.000 &   0.4 &   1.1 &   1.5 \\
  & gpt-oss-120b             &  99.9 & 0.001 &   0.8 &   2.0 &   2.3 \\
  & deepseek-r1-distill-70b  &  97.7 & 0.023 &  42.2 &  49.7 &  51.6 \\
  & phi-4                    &  95.7 & 0.043 &  66.4 &  75.5 &  78.0 \\
  & llama-3.3-70b            &  90.4 & 0.096 & 178.6 & 193.4 & 197.3 \\
  & qwq-32b                  &  89.9 & 0.101 & 183.1 & 198.2 & 202.1 \\
  & qwen3-30b-a3b            &  89.8 & 0.102 & 191.0 & 206.5 & 210.6 \\
  & qwen3-14b                &  84.0 & 0.160 & 302.7 & 321.7 & 326.7 \\
  & qwen2.5-72b              &  80.3 & 0.197 & 386.9 & 408.5 & 414.0 \\
  & granite-3.2-8b           &  70.4 & 0.296 & 588.7 & 614.7 & 621.2 \\
  & gemma-3-27b              &  68.3 & 0.317 & 625.6 & 651.9 & 659.1 \\
  & gemma-3-12b              &  61.6 & 0.384 & 760.2 & 789.1 & 796.2 \\
\midrule
\multirow{12}{*}{\rotatebox{90}{ddi}}
  & deepseek-r1-distill-70b  & 100.0 & 0.000 & 0.0 & 0.0 & 0.0 \\
  & gemma-3-12b              & 100.0 & 0.000 & 0.0 & 0.0 & 0.0 \\
  & gemma-3-27b              & 100.0 & 0.000 & 0.0 & 0.0 & 0.0 \\
  & gpt-oss-120b             & 100.0 & 0.000 & 0.0 & 0.0 & 0.0 \\
  & gpt-oss-20b              & 100.0 & 0.000 & 0.0 & 0.0 & 0.0 \\
  & llama-3.3-70b            & 100.0 & 0.000 & 0.0 & 0.0 & 0.0 \\
  & phi-4                    & 100.0 & 0.000 & 0.0 & 0.0 & 0.0 \\
  & qwen2.5-72b              & 100.0 & 0.000 & 0.0 & 0.0 & 0.0 \\
  & qwen3-14b                &  99.9 & 0.001 & 0.0 & 0.0 & 0.0 \\
  & qwq-32b                  &  99.9 & 0.001 & 0.0 & 0.0 & 0.0 \\
  & granite-3.2-8b           &  99.9 & 0.001 & 0.5 & 0.8 & 0.9 \\
  & qwen3-30b-a3b            &  99.9 & 0.001 & 1.0 & 2.1 & 2.6 \\
\bottomrule
\end{tabular}
\end{table*}

\begin{table*}[t]\centering\caption{\textbf{Insurance} per-(dataset,
model) accuracy and risk measures. $\widehat{\ES}$ and tail measures
in \$M. JudgeBERT has very heterogeneous severity profile; PrivacyQA
is near-saturation.}\label{tab:results-insurance}
\begin{tabular}{@{}llrrrrr@{}}
\toprule
Dataset & Model & Acc.\ (\%) & $\hat{p}$ & $\widehat{\ES}$ & VaR$_{95}$ & TVaR$_{95}$ \\
\midrule
\multirow{12}{*}{\rotatebox{90}{judgebert}}
  & gpt-oss-20b              & 28.5 & 0.715 & 448.8 & 464.2 & 468.4 \\
  & qwen3-14b                & 26.2 & 0.738 & 450.1 & 465.4 & 469.5 \\
  & gemma-3-27b              & 27.0 & 0.730 & 452.6 & 468.0 & 471.9 \\
  & granite-3.2-8b           & 25.1 & 0.749 & 456.7 & 472.1 & 476.3 \\
  & gemma-3-12b              & 24.7 & 0.753 & 460.4 & 476.2 & 480.2 \\
  & qwq-32b                  & 24.7 & 0.753 & 460.7 & 476.3 & 480.3 \\
  & gpt-oss-120b             & 24.9 & 0.751 & 463.7 & 479.3 & 483.3 \\
  & phi-4                    & 21.8 & 0.782 & 464.8 & 480.3 & 484.2 \\
  & qwen2.5-72b              & 23.9 & 0.761 & 465.0 & 480.6 & 484.6 \\
  & qwen3-30b-a3b            & 25.4 & 0.746 & 465.4 & 481.1 & 485.1 \\
  & llama-3.3-70b            & 19.2 & 0.808 & 476.1 & 491.8 & 495.9 \\
  & deepseek-r1-distill-70b  & 20.0 & 0.800 & 504.8 & 520.7 & 524.9 \\
\midrule
\multirow{12}{*}{\rotatebox{90}{privacyqa}}
  & gpt-oss-120b             & 99.7 & 0.003 &  0.9 &  1.6 &  1.9 \\
  & gpt-oss-20b              & 99.6 & 0.004 &  1.0 &  1.8 &  2.1 \\
  & phi-4                    & 97.1 & 0.029 &  6.1 &  8.0 &  8.6 \\
  & qwen3-14b                & 95.4 & 0.046 &  6.9 &  9.0 &  9.5 \\
  & deepseek-r1-distill-70b  & 95.9 & 0.041 &  7.2 &  9.3 &  9.9 \\
  & qwq-32b                  & 94.9 & 0.051 &  7.9 & 10.1 & 10.7 \\
  & qwen2.5-72b              & 93.5 & 0.065 & 12.6 & 15.3 & 16.1 \\
  & qwen3-30b-a3b            & 93.1 & 0.069 & 15.0 & 18.1 & 18.9 \\
  & granite-3.2-8b           & 91.0 & 0.090 & 17.5 & 20.8 & 21.7 \\
  & gemma-3-27b              & 87.5 & 0.125 & 23.9 & 27.7 & 28.7 \\
  & gemma-3-12b              & 84.6 & 0.154 & 30.1 & 34.4 & 35.5 \\
  & llama-3.3-70b            & 83.3 & 0.167 & 35.4 & 40.1 & 41.3 \\
\bottomrule
\end{tabular}
\end{table*}

\end{document}
