% =============================================================================
% Prompt Documentation for severity-eval — EMNLP 2026
% This file documents all evaluation prompts with academic references.
% Use as appendix or supplementary material.
% =============================================================================

\section{Evaluation Prompts and Methodology}
\label{sec:prompts}

We evaluate 17 LLMs across 11 datasets spanning three domains (finance, medical, legal) and two private insurance datasets.
For each dataset, we use two prompt configurations:
\begin{enumerate}
    \item \textbf{Original prompts} (primary results): zero-shot prompts from published evaluation protocols, ensuring comparability with prior work.
    \item \textbf{Standardized prompts} (ablation): uniform HELM-style templates~\cite{liang2023helm} to demonstrate that our findings are robust to prompt choice.
\end{enumerate}

All evaluations use temperature 0.0 (greedy decoding) where supported by the API, and zero-shot prompting (no in-context examples) to avoid confounding few-shot effects with severity-aware evaluation.

% ---------------------------------------------------------------------------
\subsection{Finance Domain}
% ---------------------------------------------------------------------------

\subsubsection{FinanceBench}

\textbf{Source:} Islam et al.~\cite{islam2023financebench}, Table~3 --- Oracle context-first configuration.
The original paper used human evaluation; we add a brief output constraint (``Provide only the answer'') necessary for automated scoring, following standard practice in benchmark evaluation~\cite{liang2023helm}.

\begin{quote}
\texttt{Answer this question: \{question\}} \\
\texttt{Provide only the answer, no explanation.} \\
\texttt{Context:} \\
\texttt{[START OF FILING] \{evidence\} [END OF FILING]} \\
\texttt{Answer:}
\end{quote}

Evidence is extracted from the \texttt{evidence} column of the dataset, containing SEC filing excerpts (10-K, 10-Q, 8-K) linked to each question.
Parameters: temperature 0.01, max tokens 2048 in the original paper; we use temperature 0.0, max tokens 512.

\subsubsection{FinQA}

\textbf{Source:} Chen et al.~\cite{chen2022program} (Program-of-Thoughts), confirmed by the PIXIU/FinBen benchmark~\cite{xie2024finben} (Appendix~C, Table~7).

\begin{quote}
\texttt{Read the following text and table, and then answer a question:} \\
\texttt{\{evidence\}} \\
\texttt{Question: \{question\}} \\
\texttt{Answer:}
\end{quote}

Context is assembled from the raw JSON: \texttt{pre\_text} paragraphs (joined), followed by the table (pipe-delimited rows), followed by \texttt{post\_text} paragraphs.
This matches the CoT format from~\cite{chen2022program} without the Python code generation component, as we evaluate direct answer generation.

\subsubsection{TAT-QA}

\textbf{Source:} PIXIU/FinBen benchmark~\cite{xie2024finben}, Appendix~C, Table~7; also confirmed by TAT-LLM~\cite{zhu2024tatllm}, Appendix~D.

\begin{quote}
\texttt{Read the following text and table, and then answer a question:} \\
\texttt{\{evidence\}} \\
\texttt{Question: \{question\}} \\
\texttt{Answer:}
\end{quote}

Context is assembled from the passage-level JSON: table rows (pipe-delimited) followed by paragraph texts.
The FinBen benchmark uses the same structure with the prefix ``Please answer the given financial question based on the context''~\cite{xie2024finben}; we adopt the FinQA-consistent format for uniformity across financial datasets.

% ---------------------------------------------------------------------------
\subsection{Medical Domain}
% ---------------------------------------------------------------------------

\subsubsection{MedCalc-Bench}

\textbf{Source:} NCBI evaluation code\footnote{\url{https://github.com/ncbi-nlp/MedCalc-Bench/blob/main/evaluation/run.py}} --- \texttt{direct\_answer} mode (zero-shot, no chain-of-thought).

\begin{quote}
\texttt{You are a helpful assistant for calculating a score for a given patient note.
Please output answer only without any other text.} \\[6pt]
\texttt{Here is the patient note:} \\
\texttt{\{question\}} \\[6pt]
\texttt{Please output the numerical answer:}
\end{quote}

The original code defines three modes: \texttt{direct\_answer}, \texttt{zero\_shot} (CoT), and \texttt{one\_shot}.
We use \texttt{direct\_answer} for consistency with our zero-shot evaluation protocol.
The system prompt instructs the model to output only the numerical value, matching the original evaluation methodology.

\subsubsection{MedQA USMLE}

\textbf{Source:} EleutherAI lm-evaluation-harness\footnote{\url{https://github.com/EleutherAI/lm-evaluation-harness/tree/main/lm_eval/tasks/medqa}} --- zero-shot MCQ format.

\begin{quote}
\texttt{Question: \{question\}} \\
\texttt{A. \{option\_A\}} \\
\texttt{B. \{option\_B\}} \\
\texttt{C. \{option\_C\}} \\
\texttt{D. \{option\_D\}} \\
\texttt{Answer:}
\end{quote}

This is the canonical format used across major evaluation frameworks.
HELM~\cite{liang2023helm} uses an identical structure with the prefix ``The following are multiple choice questions (with answers) about medicine'' and 5-shot examples; we omit few-shot examples for consistency.
The MedPaLM evaluation~\cite{singhal2023medpalm} uses a similar structure with chain-of-thought (``Let's think step by step''), which we exclude from the primary results.

\subsubsection{HEAD-QA}

\textbf{Source:} MedHELM~\cite{flemingmedhelm2024}\footnote{\url{https://github.com/stanford-crfm/helm/blob/main/src/helm/benchmark/run_specs/medhelm_run_specs.py}} --- zero-shot biomedical MCQ format.

\begin{quote}
\texttt{You are a highly knowledgeable AI assistant specializing in biomedical sciences.
Select the correct answer by outputting only the letter corresponding to your choice.} \\[6pt]
\texttt{Question: \{question\}} \\
\texttt{A. \{option\_A\}} \\
\texttt{...} \\
\texttt{Answer:}
\end{quote}

The biomedical instruction prefix follows MedHELM's domain-specific adapter configuration.
The original HEAD-QA paper~\cite{vilares2019headqa} evaluated IR+neural systems without LLM prompts; the lm-evaluation-harness uses a simpler format without the biomedical prefix.

% ---------------------------------------------------------------------------
\subsection{Legal Domain}
% ---------------------------------------------------------------------------

All legal prompts follow LegalBench~\cite{guha2023legalbench}, the standard benchmark for LLM evaluation on legal tasks.

\subsubsection{CUAD}

\textbf{Source:} LegalBench~\cite{guha2023legalbench} --- zero-shot jargon format.
The original CUAD paper~\cite{hendrycks2021cuad} defines an extractive QA task (SQuAD~2.0 format) where questions are the 41 clause category descriptions from \texttt{category\_descriptions.csv}.

\begin{quote}
\texttt{\{evidence\}} \\[6pt]
\texttt{Question: \{question\}} \\
\texttt{Provide only the relevant text from the clause, or `N/A' if not found.} \\
\texttt{Answer:}
\end{quote}

Evidence is the contract paragraph context from the SQuAD-format data, downloaded directly from the GitHub repository\footnote{\url{https://github.com/TheAtticusProject/cuad}} since the HuggingFace dataset script is no longer supported.

\subsubsection{MAUD}

\textbf{Source:} LegalBench~\cite{guha2023legalbench} --- MCQ classification format over merger agreement text.
MAUD~\cite{wang2023maud} defines 92 deal point questions with 2--5 answer categories per question.

\begin{quote}
\texttt{Instruction: Read the segment of a merger agreement and answer the
multiple-choice question by choosing the option that best characterizes the agreement.} \\
\texttt{Question: \{question\}} \\
\texttt{A. \{option\_A\}} \\
\texttt{B. \{option\_B\}} \\
\texttt{...} \\[6pt]
\texttt{Merger Agreement: \{evidence\}} \\
\texttt{Answer:}
\end{quote}

Answer options are pre-computed per question type from the full dataset.
The LegalBench evaluation includes both few-shot (\texttt{base\_prompt.txt}) and zero-shot (\texttt{base\_prompt\_wo\_example.txt}) variants; we use the zero-shot variant.

\subsubsection{ContractNLI}

\textbf{Source:} Adapted from LegalBench~\cite{guha2023legalbench} ContractNLI tasks.
The original paper~\cite{koreeda2021contractnli} defines a document-level NLI task with 17 fixed hypotheses over NDA contracts.
LegalBench binarizes the task (Entailment vs.\ Not Entailment); we preserve the original three-class formulation (Entailment, Contradiction, Neutral).

\begin{quote}
\texttt{Given the following NDA excerpt, determine the relationship with the hypothesis.
Answer with exactly one word: Entailment, Contradiction, or Neutral.} \\[6pt]
\texttt{Document excerpt:} \\
\texttt{\{evidence\}} \\[6pt]
\texttt{Hypothesis: \{question\}} \\
\texttt{Answer:}
\end{quote}

% ---------------------------------------------------------------------------
\subsection{Insurance Domain (Private Datasets)}
% ---------------------------------------------------------------------------

\subsubsection{Quebec Automobile Insurance RAG}

\textbf{Source:} Beauchemin et al.~\cite{beauchemin2024quebec} --- exact prompt from the RAG evaluation pipeline.

\begin{quote}
\texttt{Répondez à la question suivante sur l'assurance automobile au Québec.
Fournissez uniquement la réponse, sans explication.} \\[6pt]
\texttt{Question : \{question\}} \\
\texttt{Réponse :}
\end{quote}

The original paper uses a more detailed expert system prompt with retrieved context; we use the zero-shot variant for consistency.
Evaluation follows the manual criterion-based scoring protocol from~\cite{beauchemin2024quebec} (scale: $-1$ for false information, 0 for no answer, 1 for incomplete, 2 for complete).

\subsubsection{Legal Text Simplification (JUDGEBERT)}

\textbf{Source:} JUDGEBERT~\cite{judgebert2025} --- GPT-4 simplification prompt (French).

\begin{quote}
\texttt{Simplifiez le texte juridique suivant en langage clair,
en préservant le sens légal.} \\[6pt]
\texttt{Texte : \{question\}} \\
\texttt{Simplification :}
\end{quote}

Evaluation of simplification quality uses the Legal Meaning Preservation (LMP) score from~\cite{judgebert2025}, a 10-point Likert scale assessed by law students with brackets: 7--10 (Accurate), 2--6 (Seems Imprecise), 1 (Off-Track).

% ---------------------------------------------------------------------------
\subsection{Standardized Prompts (Ablation)}
\label{sec:standard-prompts}
% ---------------------------------------------------------------------------

For the ablation study, all datasets use a uniform HELM-style template~\cite{liang2023helm}:

\begin{quote}
\texttt{Question: \{question\}} \\
\texttt{\{context if available\}} \\
\texttt{\{options if MCQ\}} \\
\texttt{Answer:}
\end{quote}

This minimal template removes all dataset-specific instructions, domain prefixes, and formatting conventions.
If the Kendall $\tau$ between model rankings under ``original'' and ``standard'' prompts is high ($\tau > 0.8$), this confirms that our severity-weighted findings are robust to prompt engineering choices.

% ---------------------------------------------------------------------------
\subsection{Severity Annotation}
\label{sec:severity-annotation}
% ---------------------------------------------------------------------------

Each dataset instance is annotated with a severity level (critical, major, minor, negligible) using domain-specific heuristics:

\paragraph{Finance (FinanceBench, FinQA, TAT-QA).}
Two-axis classification following SEC Staff Accounting Bulletin No.~99 (SAB~99) materiality framework:
(1)~answer type (dollar-large, dollar-small, percentage, ratio, yes/no, text) $\times$
(2)~metric type (core financials, operational, ratio, descriptive).
Core financial statement items (revenue, net income, total assets) in absolute dollars are classified as \emph{critical}; operational metrics as \emph{major}; ratios and percentages as \emph{minor}; descriptive text as \emph{negligible}.

\paragraph{Medical (MedCalc, MedQA, HEAD-QA).}
Based on clinical impact: dosage/triage calculations are \emph{critical}; diagnostic/surgical decisions are \emph{major}; pathophysiology and lab interpretation are \emph{minor}; anatomy and basic science are \emph{negligible}.

\paragraph{Legal (CUAD, MAUD, ContractNLI).}
Based on clause type and legal exposure: indemnification, liability limitation, and termination clauses are \emph{critical}; IP ownership, non-compete, and deal protection are \emph{major}; governing law and closing conditions are \emph{minor}; definitions and boilerplate are \emph{negligible}.

% ---------------------------------------------------------------------------
\subsection{Scoring Methodology}
% ---------------------------------------------------------------------------

Predictions are scored using an automated cascade:
\begin{enumerate}
    \item \textbf{MCQ matching}: if options are provided, compare by option letter or exact text match.
    \item \textbf{Yes/No extraction}: word-boundary match at response start.
    \item \textbf{Numeric comparison}: extract dollar amounts (priority), percentages, or first non-year number; compare with 5\% relative tolerance.
    \item \textbf{Exact text match}: normalized (lowercase, strip punctuation, collapse whitespace).
    \item \textbf{Fuzzy containment}: reference substring in prediction (min 4 characters) or all reference words present (min 3 words).
\end{enumerate}

Number extraction prioritizes dollar amounts (\$X, \$X million/billion) over standalone numbers, and skips year-like values (1900--2099) to avoid false matches in financial documents.

% ---------------------------------------------------------------------------
% References
% ---------------------------------------------------------------------------

\begin{thebibliography}{99}

\bibitem{islam2023financebench}
P.~Islam, A.~Kannappan, D.~Kiela, R.~Qian, N.~Scherrer, and B.~Vidgen.
\newblock {FinanceBench}: A New Benchmark for Financial Question Answering.
\newblock \emph{arXiv:2311.11944}, 2023.

\bibitem{chen2022program}
W.~Chen, X.~Ma, X.~Wang, and W.~Cohen.
\newblock Program of Thoughts Prompting: Disentangling Computation from Reasoning for Numerical Reasoning Tasks.
\newblock \emph{TMLR}, 2023. arXiv:2211.12588.

\bibitem{xie2024finben}
Q.~Xie, W.~Han, Y.~Lai, M.~Peng, and J.~Huang.
\newblock {FinBen}: A Holistic Financial Benchmark for Large Language Models.
\newblock \emph{NeurIPS Datasets \& Benchmarks}, 2024. arXiv:2402.12659.

\bibitem{zhu2024tatllm}
F.~Zhu, W.~Lei, Y.~Feng, C.~Wang, B.~Zhang, and T.-S.~Chua.
\newblock {TAT-LLM}: A Specialized Language Model for Discrete Reasoning over Tabular and Textual Data.
\newblock \emph{ICAIF}, 2024. arXiv:2401.13223.

\bibitem{zhu2021tatqa}
F.~Zhu, W.~Lei, Y.~Huang, C.~Wang, S.~Zhang, J.~Lv, F.~Feng, and T.-S.~Chua.
\newblock {TAT-QA}: A Question Answering Benchmark on a Hybrid of Tabular and Textual Content in Finance.
\newblock In \emph{ACL}, pages 3277--3287, 2021.

\bibitem{liang2023helm}
P.~Liang, R.~Bommasani, T.~Lee, et~al.
\newblock Holistic Evaluation of Language Models.
\newblock \emph{TMLR}, 2023.

\bibitem{singhal2023medpalm}
K.~Singhal, S.~Azizi, T.~Tu, et~al.
\newblock Large Language Models Encode Clinical Knowledge.
\newblock \emph{Nature}, 620:172--180, 2023.

\bibitem{flemingmedhelm2024}
S.~Fleming, A.~Lozano, W.~Haberkorn, et~al.
\newblock {MedHELM}: A Medical Health Evaluation of Large Language Models.
\newblock 2024.

\bibitem{vilares2019headqa}
D.~Vilares and C.~G{\'o}mez-Rodr{\'\i}guez.
\newblock {HEAD-QA}: A Healthcare Dataset for Complex Reasoning.
\newblock In \emph{ACL}, 2019.

\bibitem{guha2023legalbench}
N.~Guha, J.~Nyarko, D.~Ho, et~al.
\newblock {LegalBench}: A Collaboratively Built Benchmark for Measuring Legal Reasoning in Large Language Models.
\newblock \emph{NeurIPS Datasets \& Benchmarks}, 2023. arXiv:2308.11462.

\bibitem{hendrycks2021cuad}
D.~Hendrycks, C.~Burns, A.~Chen, and S.~Ball.
\newblock {CUAD}: An Expert-Annotated {NLP} Dataset for Legal Contract Review.
\newblock \emph{arXiv:2103.06268}, 2021.

\bibitem{wang2023maud}
S.~Wang, A.~Holzenberger, et~al.
\newblock {MAUD}: An Expert-Annotated Legal {NLP} Dataset for Merger Agreement Understanding.
\newblock \emph{arXiv:2301.00876}, 2023.

\bibitem{koreeda2021contractnli}
Y.~Koreeda and C.~Manning.
\newblock {ContractNLI}: A Dataset for Document-level Natural Language Inference for Contracts.
\newblock In \emph{Findings of EMNLP}, pages 1907--1919, 2021.

\bibitem{beauchemin2024quebec}
D.~Beauchemin, R.~Khoury, and N.~Garneau.
\newblock Quebec Automobile Insurance Question-Answering With Retrieval-Augmented Generation.
\newblock In \emph{NLLP Workshop, ACL}, 2024.

\bibitem{judgebert2025}
JUDGEBERT: Assessing Legal Meaning Preservation Between Sentences.
\newblock In \emph{EMNLP}, 2025. arXiv:2508.16870.

\end{thebibliography}
